### 2025-05-07 Examples/Example3_3_IntegerNSynthesizerDesignedforSettlingTime/article/article.qmd
Finishing example 3.3

---
title: Integer-N PFD/CP-based PLL - System-Level simulation
subtitle: Laboratory 01 - Week 01-09
author:
  - name: Bruno Alexandre Fraga
    email: bruno.fraga@posgrad.ufsc.br
    affiliations: 
        - id: some-tech
          name: Universidade Federal de Santa Catarina
          department: Departamento de Engenharia Elétrica e Eletrônica
          address: R. Delfino Conti
          city: Florianópolis
          state: Santa Catarina
          postal-code: 88040-370
abstract: |
    Este trabalho apresenta duas simulações para um Integer-N PFD/CP-based PLL a nível de sistema. Uma simulação é no domínio do tempo e a outra no domínio da fase. Foram estabelecidas restrições para a operação e são apresentados os passos até a concepção dos sistemas para a simulação. Por fim, os resultados são analisados quanto ao tempo de resposta e comparados ao que foi previsto com base na teoria.
keywords: 
  - Phase-locked loop
  - Sistema de controle
  - PLL Integer-N
date: last-modified
bibliography:
  - bibliography.bib
  - zoterobibliography.bib
format:
  elsevier-pdf:
    include-in-header:
      - text: |
          \usepackage{steinmetz}
          \usepackage[a4paper, margin=1in]{geometry}
          \usepackage{graphicx} % Required for inserting images
          \usepackage{circuitikz}
          \usepackage{amsmath}
          \usepackage{amssymb}
          \usepackage{amsthm}
          \usepackage{comment}
          \usepackage{tikz}
          \usetikzlibrary{shapes,arrows}
          \usetikzlibrary{shapes,arrows,positioning,calc}

          \newcommand{\parallelsum}{\mathbin{\!/\mkern-5mu/\!}}
    keep-tex: true
    journal:
      name: Programa de Pós-Graduação em Engenharia Elétrica
      formatting: preprint
      # model: 3p # Don't set a model with preprint
      cite-style: number
---

```{python}
#| echo: false

from IPython.display import display, Markdown

def format(value, decimals=2):
    abs_value = abs(value)
    fmt = f".{decimals}f"

    if abs_value >= 1e9:
        result = f"{value / 1e9:{fmt}}\\,\\text{{G}}"
    elif abs_value >= 1e6:
        result = f"{value / 1e6:{fmt}}\\,\\text{{M}}"
    elif abs_value >= 1e3:
        result = f"{value / 1e3:{fmt}}\\,\\text{{k}}"
    elif abs_value >= 1:
        result = f"{value * 1e0:{fmt}}\\,"
    elif abs_value >= 1e-3:
        result = f"{value * 1e3:{fmt}}\\,\\text{{m}}"
    elif abs_value >= 1e-6:
        result = f"{value * 1e6:{fmt}}\\,\\mu"
    elif abs_value >= 1e-9:
        result = f"{value * 1e9:{fmt}}\\,\\text{{n}}"
    elif abs_value >= 1e-12:
        result = f"{value * 1e12:{fmt}}\\,\\text{{p}}"
    elif abs_value >= 1e-15:
        result = f"{value * 1e15:{fmt}}\\,\\text{{f}}"
    else:
        result = f"{value:.0e}\\,"

    return Markdown(f"${result}$")

```

```{python}
#| echo: false
import math
from IPython.display import display, Markdown

def format_with_pi_latex(value):
    pi_multiple = value / math.pi
    abs_pi_multiple = abs(pi_multiple)

    if abs_pi_multiple >= 1e9:
        result = f"{pi_multiple / 1e9:.2f}\\pi\\,\\text{{G}}"
    elif abs_pi_multiple >= 1e6:
        result = f"{pi_multiple / 1e6:.2f}\\pi\\,\\text{{M}}"
    elif abs_pi_multiple >= 1e3:
        result = f"{pi_multiple / 1e3:.2f}\\pi\\,\\text{{k}}"
    elif abs_pi_multiple >= 1:
        result = f"{pi_multiple:.2f}\\pi"
    elif abs_pi_multiple >= 1e-3:
        result = f"{pi_multiple * 1e3:.2f}\\pi\\,\\text{{m}}"
    elif abs_pi_multiple >= 1e-6:
        result = f"{pi_multiple * 1e6:.2f}\\pi\\,\\mu"
    elif abs_pi_multiple >= 1e-9:
        result = f"{pi_multiple * 1e9:.2f}\\pi\\,\\text{{n}}"
    elif abs_pi_multiple >= 1e-12:
        result = f"{pi_multiple * 1e12:.2f}\\pi\\,\\text{{p}}"
    elif abs_pi_multiple >= 1e-15:
        result = f"{pi_multiple * 1e15:.2f}\\pi\\,\\text{{f}}"
    else:
        result = f"{pi_multiple:.0e}\\pi"

    return Markdown(f"${result}$")

# Exemplo de uso:

```

```{python}
#| echo: false

f0=2.4e9
f=2.4835e9
t_s = 225e-6
f_ref = 1e6
```

# Introdução
The intention in this report is the designing of an integer-N synthesizer that operates at `{python} format(f0)`Hz to `{python} format(f)`Hz. The synthesizer must to settle for a frequency step in a settling time of, $t_s=$ `{python} format(t_s)`s. The channel spacing is `{python} format(f_ref)`Hz and as it's using an integer-N synthesizer, the channel space is equal to the frequency reference.

Using a equation discussed in the previous report, it's possible to get the natural frequency, $\omega_n$,

$$
t_s\approx\frac{-\ln\left(0.01\sqrt{1-\zeta^2}\right)}{\zeta}\frac{1}{\omega_n}
$$

```{python}
#| echo: false
import math

zeta=0.707
omega_n__t_s=math.log(0.01*math.sqrt(1-zeta**2), math.e)/(zeta)*(-1)
omega_n=omega_n__t_s/t_s
f_n=omega_n/(2*math.pi)
```

Using an arbitrary value for $\zeta=$ `{python} zeta`. With that constant, the natural frequency would be $\omega_n=$ `{python} format_with_pi_latex(omega_n)`rad/s, or $f_n=$ `{python} format(f_n)`Hz.

Starting from $omega_n$ and $zeta$, it's possible to get the 3-dB bandwidth, $\omega_{3\text{dB}}$ as follows,

$$
\omega_{3\text{dB}}=\omega_n\sqrt{1+2\zeta^2+\sqrt{4\zeta^4+4\zeta^2+2}}
$$

```{python}
#| echo: false
import math

omega_bw = omega_n * math.sqrt(1+2*zeta**2+math.sqrt(4*zeta**4+4*zeta**2+2))
f_bw = omega_bw/(2*math.pi)
```

Using that relation, it results in $f_{3\text{dB}}=$ `{python} format(f_bw)`Hz.

How the relation between $f_\text{ref}\gg f_{3\text{dB}}$, thus the system isn't affected for the discretization process (sampling). So the continuous analysis is useful for this example.

```{python}
#| echo: false
N0=f0/f_ref
N=f/f_ref
```

As it's an integer-N and the reference frequency $f_\text{ref}=$ `{python} format(f_ref)`Hz, the so $N$ should be from `{python} round(N0, 0)` to `{python} round(N, 0)`.

## VCO

```{python}
#| echo: false
import math

f_range=f-f0
tuning_range_percent = 9.6 
tuning_range = f0/tuning_range_percent
V=2
Kvco = 250e6 # Hz/V
Kvco_ang = Kvco*2*math.pi
```

To the VCO, it's needed to cover a tuning range of  `{python} format(f_range)`Hz.

Due to process variations, it is necessary to ensure that the VCO has an additional tuning margin. Therefore, a total tuning range of approximately `{python} round(tuning_range_percent, 0)`%, of the center frequency, around `{python} format(tuning_range)`Hz, is recommended. And assuming a voltage supply of `{python} format(V)`V, a VCO constant of `{python} format(Kvco)`Hz/V `{python} format(Kvco_ang)`rad/s/V is expected. So the free running frequency of the VCO should be in minimum frequency. The middle of the frequencies, $f_c$,

$$
f_c=\frac{f+f_0}{2}
$$

```{python}
#| echo: false

f_c=(f+f0)/2
fvco=f_min=f_c-tuning_range
f_max=f_c+tuning_range

f_example_1=2.401e9
```

So it results in $f_c=$ `{python} format(f_c)`Hz. The minimum frequency is $f_\text{min}=$ `{python} format(f_min)`Hz and the maximum is `{python} format(f_max)`Hz. The free running frequency will be equals to $f_\text{VCO}=f_\text{min}$. So, if it's demanded a frequency of `{python} format(f_example_1, 3)`Hz, for example, it uses the following equation to determine the voltage of control, $V_c$

$$
\begin{split}
f_o&=f_\text{VCO}+K_\text{VCO}\cdot V_c\implies\\
V_c&=\frac{f_o-f_\text{VCO}}{K_\text{VCO}}
\end{split}
$$

```{python}
#| echo: false

V_c_example_1=(f_example_1-fvco)/Kvco
```

So the voltage control will be $V_c=$ `{python} format(V_c_example_1)`V.

## Loop filter

As it was commented before, the continuous-time analysis can be useful in this system. So the already known equations for calculation of the second-order loop filter. For the first capacitor, $C_1$, its value can be obtained as

$$
C_1=\frac{I_\text{CP}K_\text{VCO}}{2\pi\cdot N\omega_n^2}
$$

```{python}
#| echo: false

import math

C_1=5e-9
I_CP_ang=C_1*2*math.pi*N0*omega_n**2/Kvco_ang
I_CP=C_1*2*math.pi*N0*omega_n**2/Kvco

```

The charge pump current, $I_\text{CP}$, is another unknown variable in this equation, so it's arbitraryly supposed a value of $C_1=$ `{python} format(C_1)`F, so the charge pump current should be $I_\text{CP}=$ `{python} format(I_CP_ang)`A. Then the resistor can be determined as

$$
R_1=2\zeta\sqrt{\frac{2\pi\cdot N}{IK_\text{VCO}C_1}}
$$

```{python}
#| echo: false

import math

R_1=2*zeta*math.sqrt((2*math.pi*N0)/(I_CP*Kvco*C_1))

```

Thus the resistor will be $R=$ `{python} format(R_1)`$\Omega$.

Using the common realtion, it's possible to find the second capacitor, $C_2$, for the second-order loop filter,

$$
C_2=\frac{C_1}{10}
$$

```{python}
#| echo: false

C_2=C_1/10 
```

Resulting in $C_2=$ `{python} format(C_2)`F.

# References {-}