### 2025-05-07 Examples/Example3_4_SimulationandEstimationofLoopSettlingTimes/article/article.qmd
Finishing report of example 3.4.

---
title: Integer-N PFD/CP-based PLL - System-Level simulation
subtitle: Laboratory 01 - Week 01-09
author:
  - name: Bruno Alexandre Fraga
    email: bruno.fraga@posgrad.ufsc.br
    affiliations: 
        - id: some-tech
          name: Universidade Federal de Santa Catarina
          department: Departamento de Engenharia Elétrica e Eletrônica
          address: R. Delfino Conti
          city: Florianópolis
          state: Santa Catarina
          postal-code: 88040-370
abstract: |
    Este trabalho apresenta duas simulações para um Integer-N PFD/CP-based PLL a nível de sistema. Uma simulação é no domínio do tempo e a outra no domínio da fase. Foram estabelecidas restrições para a operação e são apresentados os passos até a concepção dos sistemas para a simulação. Por fim, os resultados são analisados quanto ao tempo de resposta e comparados ao que foi previsto com base na teoria.
keywords: 
  - Phase-locked loop
  - Sistema de controle
  - PLL Integer-N
date: last-modified
bibliography:
  - bibliography.bib
  - zoterobibliography.bib
format:
  elsevier-pdf:
    include-in-header:
      - text: |
          \usepackage{steinmetz}
          \usepackage[a4paper, margin=1in]{geometry}
          \usepackage{graphicx} % Required for inserting images
          \usepackage{circuitikz}
          \usepackage{amsmath}
          \usepackage{amssymb}
          \usepackage{amsthm}
          \usepackage{comment}
          \usepackage{tikz}
          \usetikzlibrary{shapes,arrows}
          \usetikzlibrary{shapes,arrows,positioning,calc}

          \newcommand{\parallelsum}{\mathbin{\!/\mkern-5mu/\!}}
    keep-tex: true
    journal:
      name: Programa de Pós-Graduação em Engenharia Elétrica
      formatting: preprint
      # model: 3p # Don't set a model with preprint
      cite-style: number
---

```{python}
#| echo: false

from IPython.display import display, Markdown

def format(value, decimals=2):
    abs_value = abs(value)
    fmt = f".{decimals}f"

    if abs_value >= 1e9:
        result = f"{value / 1e9:{fmt}}\\,\\text{{G}}"
    elif abs_value >= 1e6:
        result = f"{value / 1e6:{fmt}}\\,\\text{{M}}"
    elif abs_value >= 1e3:
        result = f"{value / 1e3:{fmt}}\\,\\text{{k}}"
    elif abs_value >= 1:
        result = f"{value * 1e0:{fmt}}\\,"
    elif abs_value >= 1e-3:
        result = f"{value * 1e3:{fmt}}\\,\\text{{m}}"
    elif abs_value >= 1e-6:
        result = f"{value * 1e6:{fmt}}\\,\\mu"
    elif abs_value >= 1e-9:
        result = f"{value * 1e9:{fmt}}\\,\\text{{n}}"
    elif abs_value >= 1e-12:
        result = f"{value * 1e12:{fmt}}\\,\\text{{p}}"
    elif abs_value >= 1e-15:
        result = f"{value * 1e15:{fmt}}\\,\\text{{f}}"
    else:
        result = f"{value:.0e}\\,"

    return Markdown(f"${result}$")

```

```{python}
#| echo: false
import math
from IPython.display import display, Markdown

def format_with_pi_latex(value):
    pi_multiple = value / math.pi
    abs_pi_multiple = abs(pi_multiple)

    if abs_pi_multiple >= 1e9:
        result = f"{pi_multiple / 1e9:.2f}\\pi\\,\\text{{G}}"
    elif abs_pi_multiple >= 1e6:
        result = f"{pi_multiple / 1e6:.2f}\\pi\\,\\text{{M}}"
    elif abs_pi_multiple >= 1e3:
        result = f"{pi_multiple / 1e3:.2f}\\pi\\,\\text{{k}}"
    elif abs_pi_multiple >= 1:
        result = f"{pi_multiple:.2f}\\pi"
    elif abs_pi_multiple >= 1e-3:
        result = f"{pi_multiple * 1e3:.2f}\\pi\\,\\text{{m}}"
    elif abs_pi_multiple >= 1e-6:
        result = f"{pi_multiple * 1e6:.2f}\\pi\\,\\mu"
    elif abs_pi_multiple >= 1e-9:
        result = f"{pi_multiple * 1e9:.2f}\\pi\\,\\text{{n}}"
    elif abs_pi_multiple >= 1e-12:
        result = f"{pi_multiple * 1e12:.2f}\\pi\\,\\text{{p}}"
    elif abs_pi_multiple >= 1e-15:
        result = f"{pi_multiple * 1e15:.2f}\\pi\\,\\text{{f}}"
    else:
        result = f"{pi_multiple:.0e}\\pi"

    return Markdown(f"${result}$")

# Exemplo de uso:

```

```{python}
#| echo: false

import math

f0=3.7e9
f = 4.3e9
tunning_range = f-f0
step=1e6
fref = 40e6
I_CP=2*math.pi*100e-6
V_sup=3
f_3dB=150e3
omega_3dB=f_3dB*2*math.pi

Kvco=tunning_range/V_sup
Kvco_ang=Kvco*2*math.pi

f_30M=30e6
f_300M=300e6

Ktheta=I_CP/(2*math.pi)

fvco=4e9
N = fvco/fref
```

# Introdução
A `{python} format(f0)`Hz–`{python} format(f)`Hz synthesizer with a step size of `{python} format(step)`Hz is required. A `{python} format(fref)`Hz crystal oscillator, a charge pump with `{python} format(I_CP)`A output current, and a VCO (operating from a `{python} format(V_sup)`V supply) are available. Design a fractional-N synthesizer with a loop bandwidth of `{python} format(f_3dB)`Hz using these components. Estimate the settling time of the loop for a `{python} format(f_30M)`Hz and `{python} format(f_300M)`Hz frequency step. Simulate and compare.

## VCO

For the VCO, if the system is fed with a `{python} format(V_sup)`V supply and the tuning range is `{python} format(tunning_range)`Hz, so the VCO constant can be $K_\text{VCO}=$ `{python} format(Kvco)`Hz/V. And initially the nominal frequency is arbitrary selected as $f_\text{VCO}=$ `{python} format(fvco)`Hz. With that nominal frequency, the division ratio will be $N=$ `{python} N`.

## Charge pump

In addition, since the charge pump current is known, the charge pump constant can be determined as $K_\theta=$ `{python} format(Ktheta)`A/rad.

## Loop filter

```{python}
#| echo: false

import math

fref_f3dB=fref/f_3dB
```

As the ration of reference frequency and bandwidth is $f_\text{ref}/f_{3\text{dB}}=$ `{python} round(fref_f3dB, 0)`, i.e. $f_\text{ref}\gg f_{3\text{dB}}$, the continuous-time analysis can be used.

This section is responsible to present the design of the loop filter. First of all, it's needed to determine the value of natural frequency,

$$
\omega_n=\frac{\omega_{3\text{dB}}}{\sqrt{1+2\zeta^2+\sqrt{4\zeta^4+4\zeta^2+2}}}
$$

```{python}
#| echo: false
import math

zeta = 0.707

omega_n=omega_3dB/math.sqrt(1+2*zeta**2+math.sqrt(4*zeta**4+4*zeta**2+2))
f_n=omega_n/(2*math.pi)
```

Choosing a damping constant, $\zeta=$ `{python} zeta`, $f_n=$ `{python} format(f_n)`Hz.

The first capacitor, $C_1$, can be calculated as

$$
C_1=\frac{IK_\text{VCO}}{2\pi\cdot N\omega_n^2}
$$

```{python}
#| echo: false
import math

C_1=I_CP*Kvco_ang/(2*math.pi*N*omega_n**2)

```

So this capacitor has a value of, $C_1=$ `{python} format(C_1)`F.

The first resistor (and the single one) used in the second-order loop filter will can be obtained as follows,

$$
R_1=2\zeta\sqrt{\frac{2\pi\cdot N}{IK_\text{VCO}C_1}}
$$

```{python}
#| echo: false

import math

R_1=2*zeta*math.sqrt((2*math.pi*N)/(I_CP*Kvco_ang*C_1))
C_2=C_1/10
```

The value of that first resistor is $R_1=$ `{python} format(R_1)`$\Omega$. And the second capacitor, $C_2$, is obtained as one-tenth of C_1, getting the value of $C_2=$ `{python} format(C_2)`F.

## Testing

```{python}
#| echo: false
import math

t_m=1/(omega_n*math.sqrt(1-zeta**2))*math.atan(math.sqrt(1-zeta**2)/zeta)
```

```{python}
#| echo: false
import math

theta_e_m_n = math.sin(omega_n*math.sqrt(1-zeta**2)*t_m)/(math.sqrt(1-zeta**2))*math.e**(-zeta*omega_n*t_m)
theta_e_m = 2 * math.pi
```

```{python}
#| echo: false
import math

step_m_ang=theta_e_m*omega_n/theta_e_m_n
step_m_lin = step_m_ang/2/math.pi
```

The maximum input frequency step that can be tolerated for a system with these parameters is $\Delta\omega_m=$ `{python} format_with_pi_latex(step_m_ang)`rad/s, or `{python} format(step_m_lin)`Hz.

```{python}
#| echo: false
import math

f_i_30M = f_30M/N
f_i_30M_ang = f_i_30M*2*math.pi
f_i_300M = f_300M/N
f_i_300M_ang = f_i_300M*2*math.pi
```

```{python}
#| echo: false
import math

omega_n__t_s=math.log(0.01*math.sqrt(1-zeta**2), math.e)/(zeta)*(-1)
t_s=omega_n__t_s/omega_n
```

Now, a step in output frequency of `{python} format(f_30M)`Hz and `{python} format(f_300M)`Hz corresponds to a step in the reference frequency (at the input) of `{python} format(f_i_30M)`Hz and `{python} format(f_i_300M)`Hz, respectively. The first frequency step must be linear once it is less than the maximum frequency step, and it must be settle in a time that can be calculated using the relationship obtained for $\omega_nt_s$, so the value of settling time must be $t_s=$ `{python} format(t_s)`s.

For the step of `{python} format(f_i_300M)`Hz, in the other hand, the step will involve cycle slipping. For this nonlinear case, it's needed to estimate the acquisition time. It can be done as follows

$$
T_s=\frac{\Delta\omega}{\pi\omega_n^2}
$$

```{python}
#| echo: false
import math

Ts = f_i_300M_ang/(math.pi*omega_n**2)
Total_ts=Ts+t_s
```

Using the above equation, the acquisition time must be $T_s=$ `{python} format(Ts)`s, plus an aditional `{python} format(t_s)`s for phase lock. Therefore, the total time to phase lock, due to an step frequency of `{python} format(f_i_300M)`Hz is `{python} format(Total_ts)`s.

# References {-}