### 2025-05-14 CaseStudies/aux.qmd
askljdhfsakljhf

```{python}
#| echo: false
import math

zeta = 0.707
f_bw = 150e3
omega_bw = 2*math.pi*f_bw
```

```{python}
#| echo: false
import math

omega_n=omega_bw/math.sqrt(1+2*zeta**2+math.sqrt(4*zeta**4+4*zeta**2+2))
```

```{python}
#| echo: false
import math

t_m=1/(omega_n*math.sqrt(1-zeta**2))*math.atan(math.sqrt(1-zeta**2)/zeta)
```

```{python}
#| echo: false
import math

theta_e_m_n = math.sin(omega_n*math.sqrt(1-zeta**2)*t_m)/(math.sqrt(1-zeta**2))*math.e**(-zeta*omega_n*t_m)
theta_e_m = 2 * math.pi
```

```{python}
#| echo: false
import math

step_m_ang=theta_e_m*omega_n/theta_e_m_n
step_m_lin = step_m_ang/2/math.pi
```

```{python}
#| echo: false
import math

t_s=math.log(0.01*math.sqrt(1-zeta**2), math.e)/(zeta*omega_n)*(-1)
```