### 2025-05-28 CaseStudies/Example3_3_IntegerNSynthesizerDesignedforSettlingTime/article/article.qmd

---
title: Integer-N Frequency Synthesizer Designed for Settling Time Compliance
subtitle: Case Study - Example 3.3 from the Reference Material
author:
  - name: Bruno Alexandre Fraga
    email: bruno.fraga@posgrad.ufsc.br
    affiliations: 
        - id: some-tech
          name: Universidade Federal de Santa Catarina
          department: Departamento de Engenharia Elétrica e Eletrônica
          address: R. Delfino Conti
          city: Florianópolis
          state: Santa Catarina
          postal-code: 88040-370
abstract: |
    This report presents the design of an integer-N frequency synthesizer, focusing on the theoretical and practical aspects of system dimensioning. The objective is to ensure precise frequency generation within a specified range, while meeting constraints related to frequency resolution and settling time. The design methodology involves the definition of key dynamic parameters, such as natural frequency, damping factor, and loop bandwidth, based on continuous-time control theory. A voltage-controlled oscillator (VCO) is configured to provide the required tuning range, accounting for process variations and voltage supply limitations. Additionally, a second-order passive loop filter is designed to ensure stability and appropriate dynamic response. The overall approach demonstrates a systematic framework for the design of simple frequency synthesizers.
keywords: 
    - Integer-N PLL 
    - Settling Time
    - Continuous-Time Analysis
date: last-modified
bibliography:
  - bibliography.bib
  - zoterobibliography.bib
format:
  elsevier-pdf:
    include-in-header:
      - text: |
          \usepackage{steinmetz}
          \usepackage[a4paper, margin=1in]{geometry}
          \usepackage{graphicx} % Required for inserting images
          \usepackage{circuitikz}
          \usepackage{amsmath}
          \usepackage{amssymb}
          \usepackage{amsthm}
          \usepackage{comment}
          \usepackage{tikz}
          \usetikzlibrary{shapes,arrows}
          \usetikzlibrary{shapes,arrows,positioning,calc}

          \newcommand{\parallelsum}{\mathbin{\!/\mkern-5mu/\!}}
    keep-tex: true
    journal:
      name: Programa de Pós-Graduação em Engenharia Elétrica
      formatting: preprint
      # model: 3p # Don't set a model with preprint
      cite-style: number
---

```{python}
#| echo: false

from IPython.display import display, Markdown

def format_graph(value):
    abs_value = abs(value)
    if abs_value >= 1e9:
        result = f"{value / 1e9:.2f} G"
    elif abs_value >= 1e6:
        result = f"{value / 1e6:.2f} M"
    elif abs_value >= 1e3:
        result = f"{value / 1e3:.2f} k"
    elif abs_value >= 1:
        result = f"{value * 1e0:.2f} "
    elif abs_value >= 1e-3:
        result = f"{value * 1e3:.2f} m"
    elif abs_value >= 1e-6:
        result = f"{value * 1e6:.2f} μ"
    elif abs_value >= 1e-9:
        result = f"{value * 1e9:.2f} n"
    elif abs_value >= 1e-12:
        result = f"{value * 1e12:.2f} p"
    elif abs_value >= 1e-15:
        result = f"{value * 1e15:.2f} f"
    else:
        result = f"{value:.0e} "

    return result
```

```{python}
#| echo: false

from IPython.display import display, Markdown

def format(value, decimals=2):
    abs_value = abs(value)
    fmt = f".{decimals}f"

    if abs_value >= 1e9:
        result = f"{value / 1e9:{fmt}}\\,\\text{{G}}"
    elif abs_value >= 1e6:
        result = f"{value / 1e6:{fmt}}\\,\\text{{M}}"
    elif abs_value >= 1e3:
        result = f"{value / 1e3:{fmt}}\\,\\text{{k}}"
    elif abs_value >= 1:
        result = f"{value * 1e0:{fmt}}\\,"
    elif abs_value >= 1e-3:
        result = f"{value * 1e3:{fmt}}\\,\\text{{m}}"
    elif abs_value >= 1e-6:
        result = f"{value * 1e6:{fmt}}\\,\\mu"
    elif abs_value >= 1e-9:
        result = f"{value * 1e9:{fmt}}\\,\\text{{n}}"
    elif abs_value >= 1e-12:
        result = f"{value * 1e12:{fmt}}\\,\\text{{p}}"
    elif abs_value >= 1e-15:
        result = f"{value * 1e15:{fmt}}\\,\\text{{f}}"
    else:
        result = f"{value:.0e}\\,"

    return Markdown(f"${result}$")

```

```{python}
#| echo: false
import math
from IPython.display import display, Markdown

def format_with_pi_latex(value):
    pi_multiple = value / math.pi
    abs_pi_multiple = abs(pi_multiple)

    if abs_pi_multiple >= 1e9:
        result = f"{pi_multiple / 1e9:.2f}\\pi\\,\\text{{G}}"
    elif abs_pi_multiple >= 1e6:
        result = f"{pi_multiple / 1e6:.2f}\\pi\\,\\text{{M}}"
    elif abs_pi_multiple >= 1e3:
        result = f"{pi_multiple / 1e3:.2f}\\pi\\,\\text{{k}}"
    elif abs_pi_multiple >= 1:
        result = f"{pi_multiple:.2f}\\pi"
    elif abs_pi_multiple >= 1e-3:
        result = f"{pi_multiple * 1e3:.2f}\\pi\\,\\text{{m}}"
    elif abs_pi_multiple >= 1e-6:
        result = f"{pi_multiple * 1e6:.2f}\\pi\\,\\mu"
    elif abs_pi_multiple >= 1e-9:
        result = f"{pi_multiple * 1e9:.2f}\\pi\\,\\text{{n}}"
    elif abs_pi_multiple >= 1e-12:
        result = f"{pi_multiple * 1e12:.2f}\\pi\\,\\text{{p}}"
    elif abs_pi_multiple >= 1e-15:
        result = f"{pi_multiple * 1e15:.2f}\\pi\\,\\text{{f}}"
    else:
        result = f"{pi_multiple:.0e}\\pi"

    return Markdown(f"${result}$")

# Exemplo de uso:

```

```{python}
#| echo: false

f_o = f0=2.4e9
f=2.4835e9
t_s = 225e-6
f_r = f_ref = 1e6
```

# Introduction

This report presents the design of an integer-N frequency synthesizer operating from `{python} format(f0)`Hz to `{python} format(f, 4)`Hz. The synthesizer is required to achieve frequency settling within a time of $t_s=$ `{python} format(t_s)`s. Given that the channel spacing is `{python} format(f_ref)`Hz and an integer-N architecture is employed, the channel spacing is equal to the reference frequency. This example is based on the methodology and parameters provided in @rogers2006IntegratedCircuitDesign.

# Design Methodology

## Frequency Settling and Loop Dynamics

To meet the settling time requirement, the natural frequency $\omega_n$ of the loop must be determined. Using an equation discussed in a previous report, the relationship between settling time, damping factor $\zeta$, and natural frequency is given by:

$$
t_s\approx\frac{-\ln\left(0.01\sqrt{1-\zeta^2}\right)}{\zeta}\frac{1}{\omega_n}
$$

```{python}
#| echo: false
import math

zeta=0.707
omega_n__t_s=math.log(0.01*math.sqrt(1-zeta**2), math.e)/(zeta)*(-1)
omega_n=omega_n__t_s/t_s
f_n=omega_n/(2*math.pi)
```

Assuming an arbitrary value of $\zeta=$ `{python} zeta`, the resulting natural frequency is $\omega_n=$ `{python} format_with_pi_latex(omega_n)`rad/s, which corresponds to a frequency $f_n=$ `{python} format(f_n)`Hz.

From the natural frequency and damping factor, the 3-dB bandwidth $\omega_{3\text{dB}}$ can be calculated as:

$$
\omega_{3\text{dB}}=\omega_n\sqrt{1+2\zeta^2+\sqrt{4\zeta^4+4\zeta^2+2}}
$$

```{python}
#| echo: false
import math

omega_bw = omega_n * math.sqrt(1+2*zeta**2+math.sqrt(4*zeta**4+4*zeta**2+2))
f_bw = omega_bw/(2*math.pi)
```

This results in a bandwidth of $f_{3\text{dB}}=$ `{python} format(f_bw)`Hz. Given that $f_\text{ref}\gg f_{3\text{dB}}$, the discretization effects due to sampling are negligible, and a continuous-time analysis remains valid for this system.

```{python}
#| echo: false
N0=f0/f_ref
N=f/f_ref
```

Since the architecture is integer-N and the reference frequency is $f_\text{ref}=$ `{python} format(f_ref)`Hz, the division ratio $N$ must range from `{python} round(N0, 0)` to `{python} round(N, 0)`.

## Voltage-Controlled Oscillator (VCO)

```{python}
#| echo: false
import math

f_range=f-f0
tuning_range_percent = 9.6 
tuning_range = f0/tuning_range_percent
Vdd=2
Kvco = 250e6 # Hz/V
Kvco_ang = Kvco*2*math.pi
```

The VCO must span a tuning range of `{python} format(f_range)`Hz. To account for process variations, an additional tuning margin is considered. A total tuning range of approximately `{python} round(tuning_range_percent, 0)`% of the center frequency, or about `{python} format(tuning_range)`Hz, is recommended.

Assuming a supply voltage of `{python} format(Vdd)`V, a VCO gain of `{python} format(Kvco)`Hz/V (or `{python} format(Kvco_ang)`rad/s/V) is expected. The free-running frequency of the VCO is set at the minimum frequency. The center frequency $f_c$ is given by:

$$
f_c=\frac{f+f_0}{2}
$$

```{python}
#| echo: false

f_c=(f+f0)/2
fvco=f_min=f_c-tuning_range
f_max=f_c+tuning_range

f_example_1=2.401e9
```

This results in $f_c=$ `{python} format(f_c)`Hz. Therefore, the minimum and maximum VCO frequencies are $f_\text{min}=$ `{python} format(f_min)`Hz and $f_\text{max}=$ `{python} format(f_max)`Hz, respectively. The free-running frequency is set as $f_\text{VCO}=f_\text{min}$.

For instance, to synthesize a frequency of `{python} format(f_example_1, 3)`Hz, the required control voltage $V_c$ is computed using:

$$
\begin{split}
f_o&=f_\text{VCO}+K_\text{VCO}\cdot V_c\implies\\
V_c&=\frac{f_o-f_\text{VCO}}{K_\text{VCO}}
\end{split}
$$

```{python}
#| echo: false

V_c_example_1=(f_example_1-fvco)/Kvco
```

This yields a control voltage of $V_c=$ `{python} format(V_c_example_1)`V.

## Loop filter

As previously noted, continuous-time analysis is applicable. Standard second-order loop filter equations are used. The first capacitor $C_1$ can be calculated from:

$$
C_1=\frac{I_\text{CP}K_\text{VCO}}{2\pi\cdot N\omega_n^2}
$$

```{python}
#| echo: false

import math

C_1=5e-9
I_CP_ang=C_1*2*math.pi*N0*omega_n**2/Kvco_ang
I_CP=C_1*2*math.pi*N0*omega_n**2/Kvco

```

Since the charge pump current $I_\text{CP}$ is unknown, a value of $C_1=$ `{python} format(C_1)`F is assumed, leading to a calculated $I_\text{CP}=$ `{python} format(I_CP_ang)`A. The resistor $R_1$ is then given by:

$$
R_1=2\zeta\sqrt{\frac{2\pi\cdot N}{IK_\text{VCO}C_1}}
$$

```{python}
#| echo: false

import math

R_1=2*zeta*math.sqrt((2*math.pi*N0)/(I_CP*Kvco*C_1))

```

Thus, the resistor value is $R=$ `{python} format(R_1)`$\Omega$.

Using a common design heuristic, the second capacitor $C_2$ is:

$$
C_2=\frac{C_1}{10}
$$

```{python}
#| echo: false

C_2=C_1/10 
```

Which results in $C_2=$ `{python} format(C_2)`F.

# Parameters

## Maximum Phase Error Instant

Given the value of $\omega_n$, the time instant at which the maximum phase error occurs, $t_m$, can be computed using:

$$
t_m=\frac{1}{\omega_n\sqrt{1-\zeta^2}}\tan^{-1}\left(\frac{\sqrt{1-\zeta^2}}{\zeta}\right)
$$ {#eq-instant-maximum-frequency-step}

```{python}
#| echo: false
import math

t_m=1/(omega_n*math.sqrt(1-zeta**2))*math.atan(math.sqrt(1-zeta**2)/zeta)
```

[Eq. @eq-instant-maximum-frequency-step] represents the instant at which the phase error reaches its maximum value.

## Maximum Normalized Phase Error

The maximum normalized phase error, $\theta_{e_\text{m,n}}$, is given by:

$$
\begin{split}
\theta_{e_\text{m,n}}&=\frac{\sin\left(\omega_n\sqrt{1-\zeta^2}t_m\right)}{\sqrt{1-\zeta^2}}e^{-\zeta\omega_nt_m}
\end{split}
$$

```{python}
#| echo: false
import math

theta_e_m_n = math.sin(omega_n*math.sqrt(1-zeta**2)*t_m)/(math.sqrt(1-zeta**2))*math.e**(-zeta*omega_n*t_m)
theta_e_m = 2 * math.pi
```

And applying $t_m$ to this equation, the maximum normalized phase error in this case is approximately `{python} round(theta_e_m_n, 2)`.

## Maximum Allowed Frequency Step

This normalized value can be used in the following non-normalized equation for the phase error:

$$
\begin{split}
\theta_e(t)&=\frac{\Delta\omega}{\omega_n}\frac{\sin\left(\omega_n\sqrt{1-\zeta^2}t\right)}{\sqrt{1-\zeta^2}}e^{-\zeta\omega_nt}
\end{split}
$$

The goal is to find the maximum input frequency step, $\Delta\omega_m$, such that the maximum phase error does not exceed $\theta_{e_m}=2\pi$. Using the normalized result:

$$
\begin{split}
\theta_{e_m}&=\frac{\Delta\omega_m}{\omega_n}\theta_{e_\text{m,n}}\implies\\
\Delta\omega_m&=\frac{\theta_{e_m}\omega_n}{\theta_{e_\text{m,n}}}
\end{split}
$$

```{python}
#| echo: false
import math

step_m_ang=theta_e_m*omega_n/theta_e_m_n
step_m_lin = step_m_ang/2/math.pi
```

Substituting the calculated values yields:

- $\Delta\omega_m=$ `{python} format_with_pi_latex(step_m_ang)`rad/s
- $\Delta f_m=$ `{python} format(step_m_lin)`Hz

Hence, the maximum input frequency step that still allows theoretical predictability is `{python} format(step_m_lin)`Hz. If the input frequency step exceeds this limit, the PLL may lose lock and exhibit cycle slipping (more details are provided in the next section).

## Settling Time Estimation

The settling time, $t_s$, required for the system to reach a steady state after a frequency step can be approximated by:

$$
t_s\approx\frac{-\ln\left(0.01\sqrt{1-\zeta^2}\right)}{\zeta}\frac{1}{\omega_n}
$$

```{python}
#| echo: false
import math

t_s=math.log(0.01*math.sqrt(1-zeta**2), math.e)/(zeta*omega_n)*(-1)
```

This results in a settling time of $t_s=$ `{python} format(t_s)`s.

# Simulation

<!--```{python}
#| echo: false

import numpy as np

# Parâmetros iniciais
f_r = 10e6
f_o = 50e6
N = 5
KVCO = 10e6
I = 1e-3
Kphase = I / (2 * np.pi)
A0 = 1
K = A0 * Kphase * KVCO / N
```-->

```{python}
#| echo: false

import numpy as np

# Parâmetros iniciais
Kphase = I_CP / (2 * np.pi)
A0 = 1
K = A0 * Kphase * Kvco / N0
```

<!--```{python}
#| echo: false

from scipy import signal

omega_n = omega_bw / np.sqrt(1 + 2 * zeta**2 + np.sqrt(4 * zeta**4 + 4 * zeta**2 + 2))
f_n = omega_n/2/np.pi

# Parâmetros do PLL
C1 = K / omega_n**2
C2 = C1 / 10
R1 = 2 * zeta * omega_n / K
```-->

```{python}
#| echo: false

import math
import numpy as np
from scipy import signal

# Análise de resposta em frequência
omega_n_meas = np.sqrt(K / C_1)
f_n_meas = omega_n_meas / (2 * np.pi)

zeta_meas = (R_1 / 2) * np.sqrt(K * C_1)

t_m_meas = 1 / (omega_n_meas * math.sqrt(1 - zeta_meas**2)) * math.atan(math.sqrt(1 - zeta_meas**2) / zeta_meas)

theta_e_m_n_meas = math.sin(omega_n_meas * math.sqrt(1 - zeta_meas**2) * t_m_meas) / \
                   (math.sqrt(1 - zeta_meas**2)) * math.e**(-zeta_meas * omega_n_meas * t_m_meas)

theta_e_m_meas = 2 * math.pi

omega_bw_meas = omega_n_meas * np.sqrt(1 + 2 * zeta_meas**2 + np.sqrt(4 * zeta_meas**4 + 4 * zeta_meas**2 + 2))
f_bw_meas = omega_bw_meas / (2 * np.pi)

step_m_ang_meas = theta_e_m_meas * omega_n_meas / theta_e_m_n_meas
step_m_lin_meas = step_m_ang_meas / (2 * math.pi)

t_s_meas = math.log(0.01 * math.sqrt(1 - zeta_meas**2)) / (zeta_meas * omega_n_meas) * (-1)

```

```{python}
#| echo: false

# Numerador e denominador do compensador F(s)
# F(s) = (s*C1*R1 + 1) / (s^2*C1*C2*R1 + s*(C1 + C2))
F_num = [C_1 * R_1, 1]
F_den = [C_1 * C_2 * R_1, C_1 + C_2, 0]

F_sys = signal.TransferFunction(F_num, F_den)

# F(s) * K / s = G(s)
G_num = np.polymul(F_num, [K])
G_den = np.polymul(F_den, [1, 0])  # Multiplica por s

# H(s) = G(s) / (1 + G(s))
H_num = G_num
H_den = np.polyadd(G_den, G_num)

# Criando o sistema de tempo contínuo
H_sys = signal.TransferFunction(H_num, H_den)

```

```{python}
#| echo: false
#| label: fig-pll-step-response
#| fig-cap: "Closed-loop step response of the designed PLL."
#| fig-align: center

import numpy as np
import matplotlib.pyplot as plt
from scipy import signal

# Tempo de simulação
t = np.linspace(0, 300e-6, 100000)

# Criar vetor da entrada (degrau de amplitude 5)
step_amplitude = step_m_ang_meas
u = np.ones_like(t) * step_amplitude

# Simulação da resposta ao degrau via lsim
t_out, y_out, _ = signal.lsim(H_sys, U=u, T=t)

# Valor final da resposta
final_value = y_out[-1]
tol = 0.01 * final_value  # 1% da resposta final

# Encontrar o tempo em que a resposta entra e permanece na faixa de ±1%
ts_graphical = 0
within_bounds = np.abs(y_out - u) <= tol
for i in range(len(within_bounds)):
    if all(within_bounds[i:]):
        ts_graphical = t_out[i]
        break

# Gerar gráfico
plt.figure(figsize=(8, 5))
plt.plot(t_out * 1e6, y_out, label='Step Response')
plt.plot(t_out * 1e6, u, linestyle=':', color='black', label='Step')
plt.axhline(final_value + tol, color='red', linestyle='--', linewidth=0.8, label='±1% Tolerance')
plt.axhline(final_value - tol, color='red', linestyle='--', linewidth=0.8)
plt.axvline(ts_graphical * 1e6, color='green', linestyle='--', linewidth=1.2, label=f'Graphical settling time: {ts_graphical*1e6:.5f} us')
plt.xlabel('Time [us]')
plt.ylabel('Phase [rad]')
plt.title('Step Response with Settling Time (99%)')
plt.grid(True)
plt.legend()
plt.tight_layout()
plt.show()
```

Using the parameters $\omega_n$ and $\zeta$, the transfer function of the entire loop can be described as (neglecting the second capacitor, $C_2$, of the filter):

$$
\frac{\theta_o}{\theta_R}=\frac{\omega_n^2\left(\frac{2\zeta}{\omega_n}s+1\right)}{s^2+2\zeta\omega_n s+\omega_n^2}
$$ {#eq-full-transfer-function-no-C2}

The natural frequency of the loop, $\omega_n$, was computed from the loop bandwidth $\omega_\mathrm{bw}$ and damping factor $\zeta$, yielding a natural frequency $f_n=\omega_n/2\pi=$ `{python} format(f_n)`Hz.

Based on the chosen parameters, the loop filter components were calculated as follows. The primary capacitor of the loop filter is:

$$
C_1 = \frac{K_\theta K_\mathrm{VCO}}{N \omega_n^2}
$$

That is, $C_1=$ `{python} format(C_1)`F.

The second capacitor is chosen to be one-tenth of $C_1$, i.e., $C_2=$ `{python} format(C_2)`F.

The resistor of the loop filter is given by:

$$
R_1 = \frac{\zeta 4 \pi N \omega_n}{I K_\mathrm{VCO}}
$$

That is, $R_1=$ `{python} format(R_1)`$\Omega$.

These component values define the loop filter required to achieve the desired dynamic behavior of the PLL. With these components, it's possible to extract another set of the system parameters, i.e., $f_n=\omega_n/2\pi$, $\zeta$ and $f_\mathrm{bw}=\omega_\mathrm{bw}/2\pi$. The values are $f_n=$ `{python} format(f_n_meas)`Hz, $\zeta=$ `{python} round(zeta_meas, 3)` and $f_\mathrm{bw}=$ `{python} format(f_bw_meas)`Hz.

Similarlly what was presented in [Eq. @eq-full-transfer-function-no-C2], it's possible to present the transfer function of the system in terms of components values. This transfer function can be seen below (this equation uses the second capacitor, $C_2$, of the loop filter, adding a high-frequency pole to the system):

$$
\begin{split}
\frac{\theta_o}{\theta_R}&=\frac{K_\mathrm{VCO}K_\theta\left(1+sC_1R\right)}{s^2N\left(C_1+C_2\right)\left(1+sC_sR\right)+K_\mathrm{VCO}K_\theta\left(1+sC_1R\right)}
\end{split}
$$

where $C_s$ is the paralel between $C_1$ and $C_2$. With the PLL values properly defined, the transfer function of this system is given by:

```{python}
#| echo: false

from scipy import signal
from IPython.display import display, Markdown

def format_eng(x):
    if x == 0:
        return "0"
    exponent = int("{:.0e}".format(x).split('e')[1])
    base = x / (10**exponent)
    return f"{base:.2f}e{exponent}"

def poly_to_str(coeffs):
    terms = []
    degree = len(coeffs) - 1
    for i, coef in enumerate(coeffs):
        power = degree - i
        if abs(coef) < 1e-12:  # ignora coeficientes muito pequenos
            continue
        term = ""
        coef_str = format_eng(coef)
        if not (abs(coef - 1) < 1e-12 and power != 0):
            term += coef_str
        if power > 0:
            term += "s"
            if power > 1:
                term += f"^{power}"
        terms.append(term)
    return " + ".join(terms)

# Em scipy.signal, num e den são arrays dentro de listas (mesmo para SISO)
num = H_sys.num[0][0] if isinstance(H_sys.num[0], (list, tuple)) else H_sys.num
den = H_sys.den[0][0] if isinstance(H_sys.den[0], (list, tuple)) else H_sys.den

num_str = poly_to_str(num)
den_str = poly_to_str(den)

# Exibe como equação no Jupyter
display(Markdown(f"$$H(s) = \\frac{{{num_str}}}{{{den_str}}}$$ {{#eq-transfer-function}}"))

```

```{python}
#| echo: false

from scipy import signal
from IPython.display import display, Markdown

def format_eng(x):
    if x == 0:
        return "0"
    exponent = int("{:.0e}".format(x).split('e')[1])
    base = x / (10**exponent)
    return f"{base:.2f}e{exponent}"

def poly_to_str(coeffs):
    terms = []
    degree = len(coeffs) - 1
    for i, coef in enumerate(coeffs):
        power = degree - i
        if abs(coef) < 1e-12:  # ignora coeficientes muito pequenos
            continue
        term = ""
        coef_str = format_eng(coef)
        if not (abs(coef - 1) < 1e-12 and power != 0):
            term += coef_str
        if power > 0:
            term += "s"
            if power > 1:
                term += f"^{power}"
        terms.append(term)
    return " + ".join(terms)

# Em scipy.signal, num e den são arrays dentro de listas (mesmo para SISO)
num = F_sys.num[0][0] if isinstance(F_sys.num[0], (list, tuple)) else F_sys.num
den = F_sys.den[0][0] if isinstance(F_sys.den[0], (list, tuple)) else F_sys.den

num_str = poly_to_str(num)
den_str = poly_to_str(den)

# Exibe como equação no Jupyter
display(Markdown(f"$$F(s) = \\frac{{{num_str}}}{{{den_str}}}$$ {{#eq-transfer-function-loop-filter}}"))

```

A step with the same amplitude as the previously calculated frequency step was also applied in phase, that is, $\Delta\theta=$ `{python} format_with_pi_latex(step_m_ang_meas)`rad/s. Applying this signal to [Eq. @eq-transfer-function], the response of this system to the phase step is obtained. With this step, the settling time achieved is $t_s=$ `{python} format(t_s_meas)`s. Through simulation, it is possible to verify that the obtained value is $t_s=$ `{python} format(ts_graphical)`s. The simulated response can be visualized in [Fig. @fig-pll-step-response].

```{python}
#| echo: false

import numpy as np
import matplotlib.pyplot as plt
from scipy import signal

# Gerar a resposta em frequência
w, mag, phase = signal.bode(H_sys)

# Encontrar o ponto onde a magnitude cruza -3 dB
cutoff_index = np.argmax(mag <= -3)
cutoff_freq_ang = w[cutoff_index]
cutoff_freq_lin = cutoff_freq_ang/2/np.pi
cutoff_mag = mag[cutoff_index]

# Plotando o diagrama de Bode
plt.figure(figsize=(10, 6))

# Magnitude
plt.subplot(2, 1, 1)
plt.semilogx(w, mag, label='Magnitude (dB)')
plt.axvline(x=cutoff_freq_ang, color='r', linestyle='--', label='Cutoff Frequency (-3 dB)')
plt.axhline(y=-3, color='gray', linestyle=':', label='-3 dB')
plt.scatter(cutoff_freq_ang, cutoff_mag, color='red')  # Marcar o ponto

# Adiciona o valor da frequência no gráfico
plt.text(cutoff_freq_ang, cutoff_mag - 5, f'{format_graph(cutoff_freq_lin)}Hz', color='red',
         ha='center', va='top', fontsize=9, bbox=dict(facecolor='white', edgecolor='red', boxstyle='round'))

plt.title('Frequency Response - Bode Plot')
plt.ylabel('Magnitude (dB)')
plt.grid(True, which="both", ls="--")
plt.legend()

# Fase
plt.subplot(2, 1, 2)
plt.semilogx(w, phase)
plt.axvline(x=cutoff_freq_ang, color='r', linestyle='--')
plt.ylabel('Phase (degrees)')
plt.xlabel('Frequency (rad/s)')
plt.grid(True, which="both", ls="--")

plt.tight_layout()
plt.show()
```

```{python}
#| echo: false

import numpy as np
import matplotlib.pyplot as plt
from scipy import signal

# Gerar a resposta em frequência
w, mag, phase = signal.bode(F_sys)

# Encontrar o ponto onde a magnitude cruza -3 dB
cutoff_index = np.argmax(mag <= -3)
cutoff_freq_ang = w[cutoff_index]
cutoff_freq_lin = cutoff_freq_ang/2/np.pi
cutoff_mag = mag[cutoff_index]

# Plotando o diagrama de Bode
plt.figure(figsize=(10, 6))

# Magnitude
plt.subplot(2, 1, 1)
plt.semilogx(w, mag, label='Magnitude (dB)')

plt.title('Bode Plot - Loop filter')
plt.ylabel('Magnitude (dB)')
plt.grid(True, which="both", ls="--")
plt.legend()

# Fase
plt.subplot(2, 1, 2)
plt.semilogx(w, phase)
plt.ylabel('Phase (degrees)')
plt.xlabel('Frequency (rad/s)')
plt.grid(True, which="both", ls="--")

plt.tight_layout()
plt.show()
```

## Cycle slipping analyzing

```{python}
#| echo: false

import numpy as np

omega_r_safe = omega_n*10
f_r_safe = omega_r_safe/2/np.pi

omega_r_max = omega_n * 2 * np.pi * zeta
f_r_max = omega_r_max/2/np.pi
```

With the intention of conducting tests to verify the occurrence of cycle slipping, the minimum frequency of the reference signal is estimated. A ratio often quoted as being safe is 10:1 for $\omega_r/\omega_n$; thus, the minimum safe frequency for the reference signal is given by $f_r=$ `{python} format(f_r_safe)`Hz. The minimum allowed frequency is given by:

$$
\frac{\omega_r}{\omega_n}\ge 2\pi\zeta
$$

That is, the minimum reference frequency is $f_r\ge$ `{python} format(f_r_max)`Hz. Therefore, since a PLL operating with a reference frequency of $f_r=$ `{python} format(f_r)`Hz and an output frequency of `{python} format(f_o)`Hz is desired, the designed system must operate correctly.

Using the Simulink software, a testbench was prepared to verify the settling time of the PLL. With the previously calculated parameters and using the maximum step frequency value, the resulting curve can be seen in [Fig. @fig-max-step-response]. It can be observed that the step is applied at the instant of 1 ms, as the system is already properly stabilized at that moment. As predicted, cycle slipping does not occur; however, unlike what was calculated, the settling time of this signal is $9.196\;\mu\text{s}$.

<!--![System response to the maximum step frequency](../images/max_step_frequency_response.jpg){#fig-max-step-response}-->

```{python}
#| echo: false

cycle_slip_step_lin = 3*step_m_lin
```

Moreover, contrary to what was previously predicted, for cycle slipping to occur, the step frequency must be approximately $\Delta f=$ `{python} format(cycle_slip_step_lin)`Hz. The system's response to this signal can be seen in [Fig. @fig-cycle-slip-response].

<!--![System response with cycle slip](../images/step_frequency_response_cycle_slip.jpg){#fig-cycle-slip-response}-->

# Conclusion

This report has outlined the key steps in designing an integer-N frequency synthesizer operating within the specified frequency range. A continuous-time model was adopted due to the large separation between the reference frequency and the loop bandwidth. Component values for the VCO and loop filter were estimated based on reasonable assumptions, providing a solid foundation for subsequent implementation and simulation.

# References {-}