### 2025-06-01 CaseStudies/Example3_3_IntegerNSynthesizerDesignedforSettlingTime/article/article.qmd
Finishing charts of example 3.3.

---
title: Integer-N Frequency Synthesizer Designed for Settling Time Compliance
subtitle: Case Study - Example 3.3 from the Reference Material
author:
  - name: Bruno Alexandre Fraga
    email: bruno.fraga@posgrad.ufsc.br
    affiliations: 
        - id: some-tech
          name: Universidade Federal de Santa Catarina
          department: Departamento de Engenharia Elétrica e Eletrônica
          address: R. Delfino Conti
          city: Florianópolis
          state: Santa Catarina
          postal-code: 88040-370
abstract: |
    This report presents the design of an integer-N frequency synthesizer, focusing on the theoretical and practical aspects of system dimensioning. The objective is to ensure precise frequency generation within a specified range, while meeting constraints related to frequency resolution and settling time. The design methodology involves the definition of key dynamic parameters, such as natural frequency, damping factor, and loop bandwidth, based on continuous-time control theory. A voltage-controlled oscillator (VCO) is configured to provide the required tuning range, accounting for process variations and voltage supply limitations. Additionally, a second-order passive loop filter is designed to ensure stability and appropriate dynamic response. The overall approach demonstrates a systematic framework for the design of simple frequency synthesizers.
keywords: 
    - Integer-N PLL 
    - Settling Time
    - Continuous-Time Analysis
date: last-modified
bibliography:
  - bibliography.bib
  - zoterobibliography.bib
format:
  elsevier-pdf:
    include-in-header:
      - text: |
          \usepackage{steinmetz}
          \usepackage[a4paper, margin=1in]{geometry}
          \usepackage{graphicx} % Required for inserting images
          \usepackage{circuitikz}
          \usepackage{amsmath}
          \usepackage{amssymb}
          \usepackage{amsthm}
          \usepackage{comment}
          \usepackage{tikz}
          \usetikzlibrary{shapes,arrows}
          \usetikzlibrary{shapes,arrows,positioning,calc}

          \newcommand{\parallelsum}{\mathbin{\!/\mkern-5mu/\!}}
    keep-tex: true
    journal:
      name: Programa de Pós-Graduação em Engenharia Elétrica
      formatting: preprint
      # model: 3p # Don't set a model with preprint
      cite-style: number
---

```{python}
#| echo: false

from IPython.display import display, Markdown

def format_graph(value):
    abs_value = abs(value)
    if abs_value >= 1e9:
        result = f"{value / 1e9:.2f} G"
    elif abs_value >= 1e6:
        result = f"{value / 1e6:.2f} M"
    elif abs_value >= 1e3:
        result = f"{value / 1e3:.2f} k"
    elif abs_value >= 1:
        result = f"{value * 1e0:.2f} "
    elif abs_value >= 1e-3:
        result = f"{value * 1e3:.2f} m"
    elif abs_value >= 1e-6:
        result = f"{value * 1e6:.2f} μ"
    elif abs_value >= 1e-9:
        result = f"{value * 1e9:.2f} n"
    elif abs_value >= 1e-12:
        result = f"{value * 1e12:.2f} p"
    elif abs_value >= 1e-15:
        result = f"{value * 1e15:.2f} f"
    else:
        result = f"{value:.0e} "

    return result
```

```{python}
#| echo: false

from IPython.display import display, Markdown

def format(value, decimals=2):
    abs_value = abs(value)
    fmt = f".{decimals}f"

    if abs_value >= 1e9:
        result = f"{value / 1e9:{fmt}}\\,\\text{{G}}"
    elif abs_value >= 1e6:
        result = f"{value / 1e6:{fmt}}\\,\\text{{M}}"
    elif abs_value >= 1e3:
        result = f"{value / 1e3:{fmt}}\\,\\text{{k}}"
    elif abs_value >= 1:
        result = f"{value * 1e0:{fmt}}\\,"
    elif abs_value >= 1e-3:
        result = f"{value * 1e3:{fmt}}\\,\\text{{m}}"
    elif abs_value >= 1e-6:
        result = f"{value * 1e6:{fmt}}\\,\\mu"
    elif abs_value >= 1e-9:
        result = f"{value * 1e9:{fmt}}\\,\\text{{n}}"
    elif abs_value >= 1e-12:
        result = f"{value * 1e12:{fmt}}\\,\\text{{p}}"
    elif abs_value >= 1e-15:
        result = f"{value * 1e15:{fmt}}\\,\\text{{f}}"
    else:
        result = f"{value:.0e}\\,"

    return Markdown(f"${result}$")

```

```{python}
#| echo: false
import math
from IPython.display import display, Markdown

def format_with_pi_latex(value):
    pi_multiple = value / math.pi
    abs_pi_multiple = abs(pi_multiple)

    if abs_pi_multiple >= 1e9:
        result = f"{pi_multiple / 1e9:.2f}\\pi\\,\\text{{G}}"
    elif abs_pi_multiple >= 1e6:
        result = f"{pi_multiple / 1e6:.2f}\\pi\\,\\text{{M}}"
    elif abs_pi_multiple >= 1e3:
        result = f"{pi_multiple / 1e3:.2f}\\pi\\,\\text{{k}}"
    elif abs_pi_multiple >= 1:
        result = f"{pi_multiple:.2f}\\pi"
    elif abs_pi_multiple >= 1e-3:
        result = f"{pi_multiple * 1e3:.2f}\\pi\\,\\text{{m}}"
    elif abs_pi_multiple >= 1e-6:
        result = f"{pi_multiple * 1e6:.2f}\\pi\\,\\mu"
    elif abs_pi_multiple >= 1e-9:
        result = f"{pi_multiple * 1e9:.2f}\\pi\\,\\text{{n}}"
    elif abs_pi_multiple >= 1e-12:
        result = f"{pi_multiple * 1e12:.2f}\\pi\\,\\text{{p}}"
    elif abs_pi_multiple >= 1e-15:
        result = f"{pi_multiple * 1e15:.2f}\\pi\\,\\text{{f}}"
    else:
        result = f"{pi_multiple:.0e}\\pi"

    return Markdown(f"${result}$")

# Exemplo de uso:

```

```{python}
#| echo: false

import numpy as np

f_o = f0=2.4e9
omega_0 = 2 * np.pi * f0
f=2.4835e9
omega_f = np.pi * 2 * f
t_s = 225e-6
f_r = f_ref = 1e6
omega_r = np.pi * 2 * f_r
```

# Introduction

This report presents the design of an integer-N frequency synthesizer operating from `{python} format(f0)`Hz to `{python} format(f, 4)`Hz. The synthesizer is required to achieve frequency settling within a time of $t_s=$ `{python} format(t_s)`s. Given that the channel spacing is `{python} format(f_ref)`Hz and an integer-N architecture is employed, the channel spacing is equal to the reference frequency. This example is based on the methodology and parameters provided in @rogers2006IntegratedCircuitDesign.

# System Modeling

## Frequency Settling and Loop Dynamics

To meet the settling time requirement, the natural frequency $\omega_n$ of the loop must be determined. Using an equation discussed in a previous report, the relationship between settling time, damping factor $\zeta$, and natural frequency is given by:

$$
t_s\approx\frac{-\ln\left(0.01\sqrt{1-\zeta^2}\right)}{\zeta}\frac{1}{\omega_n}
$$

```{python}
#| echo: false
import math

zeta=0.707
omega_n__t_s=math.log(0.01*math.sqrt(1-zeta**2), math.e)/(zeta)*(-1)
omega_n=omega_n__t_s/t_s
f_n=omega_n/(2*math.pi)
```

Assuming an arbitrary value of $\zeta=$ `{python} zeta`, the resulting natural frequency is $\omega_n=$ `{python} format_with_pi_latex(omega_n)`rad/s, which corresponds to a frequency $f_n=$ `{python} format(f_n)`Hz.

From the natural frequency and damping factor, the 3-dB bandwidth $\omega_{3\text{dB}}$ can be calculated as:

$$
\omega_{3\text{dB}}=\omega_n\sqrt{1+2\zeta^2+\sqrt{4\zeta^4+4\zeta^2+2}}
$$

```{python}
#| echo: false
import math

omega_bw = omega_n * math.sqrt(1+2*zeta**2+math.sqrt(4*zeta**4+4*zeta**2+2))
f_bw = omega_bw/(2*math.pi)
```

This results in a bandwidth of $f_{3\text{dB}}=$ `{python} format(f_bw)`Hz. Given that $f_\text{ref}\gg f_{3\text{dB}}$, the discretization effects due to sampling are negligible, and a continuous-time analysis remains valid for this system.

```{python}
#| echo: false
N0=f0/f_ref
N=f/f_ref
```

Since the architecture is integer-N and the reference frequency is $f_\text{ref}=$ `{python} format(f_ref)`Hz, the division ratio $N$ must range from `{python} round(N0, 0)` to `{python} round(N, 0)`.

## Voltage-Controlled Oscillator (VCO)

```{python}
#| echo: false
import math

f_range=f-f0
tuning_range_percent = 9.6 
tuning_range = f0/tuning_range_percent
Vdd=2
KVCO = Kvco = 250e6 # Hz/V
KVCO_ang = Kvco_ang = Kvco*2*math.pi
```

The VCO must span a tuning range of `{python} format(f_range)`Hz. To account for process variations, an additional tuning margin is considered. A total tuning range of approximately `{python} round(tuning_range_percent, 0)`% of the center frequency, or about `{python} format(tuning_range)`Hz, is recommended.

Assuming a supply voltage of `{python} format(Vdd)`V, a VCO gain of `{python} format(Kvco)`Hz/V (or `{python} format(Kvco_ang)`rad/s/V) is expected. The free-running frequency of the VCO is set at the minimum frequency. The center frequency $f_c$ is given by:

$$
f_c=\frac{f+f_0}{2}
$$

```{python}
#| echo: false

f_c=(f+f0)/2
fvco=f_min=f_c-tuning_range
f_max=f_c+tuning_range

f_example_1=2.401e9
```

This results in $f_c=$ `{python} format(f_c)`Hz. Therefore, the minimum and maximum VCO frequencies are $f_\text{min}=$ `{python} format(f_min)`Hz and $f_\text{max}=$ `{python} format(f_max)`Hz, respectively. The free-running frequency is set as $f_\text{VCO}=f_\text{min}$.

For instance, to synthesize a frequency of `{python} format(f_example_1, 3)`Hz, the required control voltage $V_c$ is computed using:

$$
\begin{split}
f_o&=f_\text{VCO}+K_\text{VCO}\cdot V_c\implies\\
V_c&=\frac{f_o-f_\text{VCO}}{K_\text{VCO}}
\end{split}
$$

```{python}
#| echo: false

V_c_example_1=(f_example_1-fvco)/Kvco
```

This yields a control voltage of $V_c=$ `{python} format(V_c_example_1)`V.

## Loop filter

As previously noted, continuous-time analysis is applicable. Standard second-order loop filter equations are used. The first capacitor $C_1$ can be calculated from:

```{python}
#| echo: false

import numpy as np

# Parâmetros iniciais
#K_theta = Kphase = I_CP / (2 * np.pi)
A0 = 6.258
#A0 = 1
#K = A0 * Kphase * Kvco / N0
```

```{python}
#| echo: false

# Parâmetros do PLL
#C1 = K / omega_n**2
#C2 = C1 / 10
#R1 = 2 * zeta * omega_n / K
```

$$
C_1=\frac{I_\text{CP}K_\text{VCO}}{2\pi\cdot N\omega_n^2}
$$

```{python}
#| echo: false

import math 
import numpy as np

C_1=C1=5e-9
I_CP_ang=C_1*2*math.pi*N0*omega_n**2/Kvco_ang/A0
I = I_CP=C_1*2*math.pi*N0*omega_n**2/Kvco/A0
K_theta = Kphase = I_CP / (2 * np.pi)

K = A0 * Kphase * Kvco / N0
#K_theta = I_CP/2/math.pi
```

Since the charge pump current $I_\text{CP}$ is unknown, a value of $C_1=$ `{python} format(C_1)`F is assumed, leading to a calculated $I_\text{CP}=$ `{python} format(I_CP_ang)`A. The resistor $R_1$ is then given by:

$$
R_1=2\zeta\sqrt{\frac{2\pi\cdot N}{IK_\text{VCO}C_1}}
$$

```{python}
#| echo: false

import math

R_1=R1 = 2 * zeta * omega_n / K

```

Thus, the resistor value is $R=$ `{python} format(R_1)`$\Omega$.

Using a common design heuristic, the second capacitor $C_2$ is:

$$
C_2=\frac{C_1}{10}
$$

```{python}
#| echo: false

C_2=C2=C_1/10 
```

Which results in $C_2=$ `{python} format(C_2)`F.

# Design Analysis (wrong)

## Maximum Phase Error Instant

Given the value of $\omega_n$, the time instant at which the maximum phase error occurs, $t_m$, can be computed using:

$$
t_m=\frac{1}{\omega_n\sqrt{1-\zeta^2}}\tan^{-1}\left(\frac{\sqrt{1-\zeta^2}}{\zeta}\right)
$$ {#eq-instant-maximum-frequency-step}

```{python}
#| echo: false
import math

t_m=1/(omega_n*math.sqrt(1-zeta**2))*math.atan(math.sqrt(1-zeta**2)/zeta)
```

[Eq. @eq-instant-maximum-frequency-step] represents the instant at which the phase error reaches its maximum value.

## Maximum Normalized Phase Error

The maximum normalized phase error, $\theta_{e_\text{m,n}}$, is given by:

$$
\begin{split}
\theta_{e_\text{m,n}}&=\frac{\sin\left(\omega_n\sqrt{1-\zeta^2}t_m\right)}{\sqrt{1-\zeta^2}}e^{-\zeta\omega_nt_m}
\end{split}
$$

```{python}
#| echo: false
import math

theta_e_m_n = math.sin(omega_n*math.sqrt(1-zeta**2)*t_m)/(math.sqrt(1-zeta**2))*math.e**(-zeta*omega_n*t_m)
theta_e_m = 1.6255 * math.pi
```

And applying $t_m$ to this equation, the maximum normalized phase error in this case is approximately `{python} round(theta_e_m_n, 2)`.

## Maximum Allowed Frequency Step

This normalized value can be used in the following non-normalized equation for the phase error:

$$
\begin{split}
\theta_e(t)&=\frac{\Delta\omega}{\omega_n}\frac{\sin\left(\omega_n\sqrt{1-\zeta^2}t\right)}{\sqrt{1-\zeta^2}}e^{-\zeta\omega_nt}
\end{split}
$$

The goal is to find the maximum input frequency step, $\Delta\omega_m$, such that the maximum phase error does not exceed $\theta_{e_m}=2\pi$. Using the normalized result:

$$
\begin{split}
\theta_{e_m}&=\frac{\Delta\omega_m}{\omega_n}\theta_{e_\text{m,n}}\implies\\
\Delta\omega_m&=\frac{\theta_{e_m}\omega_n}{\theta_{e_\text{m,n}}}
\end{split}
$$

```{python}
#| echo: false
import math

step_m_ang=theta_e_m*omega_n/theta_e_m_n
step_m_lin = step_m_ang/2/math.pi
```

Substituting the calculated values yields:

- $\Delta\omega_m=$ `{python} format_with_pi_latex(step_m_ang)`rad/s
- $\Delta f_m=$ `{python} format(step_m_lin)`Hz

Hence, the maximum input frequency step that still allows theoretical predictability is `{python} format(step_m_lin)`Hz. If the input frequency step exceeds this limit, the PLL may lose lock and exhibit cycle slipping (more details are provided in the next section).

## Settling Time Estimation

The settling time, $t_s$, required for the system to reach a steady state after a frequency step can be approximated by:

$$
t_s\approx\frac{-\ln\left(0.01\sqrt{1-\zeta^2}\right)}{\zeta}\frac{1}{\omega_n}
$$

```{python}
#| echo: false
import math

t_s=math.log(0.01*math.sqrt(1-zeta**2), math.e)/(zeta*omega_n)*(-1)
```

This results in a settling time of $t_s=$ `{python} format(t_s)`s.

# System Responses

```{python}
#| echo: false

import math
import numpy as np
from scipy import signal

# Análise de resposta em frequência
omega_n_meas = np.sqrt(K / C_1)
f_n_meas = omega_n_meas / (2 * np.pi)

zeta_meas = (R_1 / 2) * np.sqrt(K * C_1)

t_m_meas = 1 / (omega_n_meas * math.sqrt(1 - zeta_meas**2)) * math.atan(math.sqrt(1 - zeta_meas**2) / zeta_meas)

theta_e_m_n_meas = math.sin(omega_n_meas * math.sqrt(1 - zeta_meas**2) * t_m_meas) / \
                   (math.sqrt(1 - zeta_meas**2)) * math.e**(-zeta_meas * omega_n_meas * t_m_meas)

theta_e_m_meas = theta_e_m

omega_bw_meas = omega_n_meas * np.sqrt(1 + 2 * zeta_meas**2 + np.sqrt(4 * zeta_meas**4 + 4 * zeta_meas**2 + 2))
f_bw_meas = omega_bw_meas / (2 * np.pi)

step_m_ang_meas = theta_e_m_meas * omega_n_meas / theta_e_m_n_meas
step_m_lin_meas = step_m_ang_meas / (2 * math.pi)

t_s_meas = math.log(0.01 * math.sqrt(1 - zeta_meas**2)) / (zeta_meas * omega_n_meas) * (-1)
```

```{python}
#| echo: false

from scipy import signal
import numpy as np

# Numerador e denominador do compensador F(s)
# F(s) = (s*C1*R1 + 1) / (s^2*C1*C2*R1 + s*(C1 + C2))
F_num = [C_1 * R_1, 1]
F_den = [C_1 * C_2 * R_1, C_1 + C_2, 0]

F_sys = signal.TransferFunction(F_num, F_den)

# F(s) * K / s = G(s)
G_num = np.polymul(F_num, [K])
G_den = np.polymul(F_den, [1, 0])  # Multiplica por s

# H(s) = G(s) / (1 + G(s))
H_num = G_num
H_den = np.polyadd(G_den, G_num)

# Criando o sistema de tempo contínuo
H_sys = signal.TransferFunction(H_num, H_den)

```

```{python}
#| echo: false
#| label: fig-pll-step-response-lalalala
#| fig-cap: "Closed-loop step response of the designed PLL."
#| fig-align: center

import numpy as np
import matplotlib.pyplot as plt
from scipy import signal

# Tempo de simulação
t = np.linspace(0, 300e-6, 100000)

# Criar vetor da entrada (degrau de amplitude 5)
step_amplitude = step_m_ang_meas
u = np.ones_like(t) * step_amplitude

# Simulação da resposta ao degrau via lsim
t_out, y_out, _ = signal.lsim(H_sys, U=u, T=t)

# Valor final da resposta
final_value = y_out[-1]
tol = 0.01 * final_value  # 1% da resposta final

# Encontrar o tempo em que a resposta entra e permanece na faixa de ±1%
ts_graphical = 0
within_bounds = np.abs(y_out - u) <= tol
for i in range(len(within_bounds)):
    if all(within_bounds[i:]):
        ts_graphical = t_out[i]
        break

# Gerar gráfico
plt.figure(figsize=(8, 5))
plt.plot(t_out * 1e6, y_out, label='Step Response')
plt.plot(t_out * 1e6, u, linestyle=':', color='black', label='Step')
plt.axhline(final_value + tol, color='red', linestyle='--', linewidth=0.8, label='±1% Tolerance')
plt.axhline(final_value - tol, color='red', linestyle='--', linewidth=0.8)
plt.axvline(ts_graphical * 1e6, color='green', linestyle='--', linewidth=1.2, label=f'Graphical settling time: {ts_graphical*1e6:.5f} us')
plt.xlabel('Time [us]')
plt.ylabel('Phase [rad]')
plt.title('Step Response with Settling Time (99%)')
plt.grid(True)
plt.legend()
plt.tight_layout()
plt.show()
```

```{python}
#| echo: false
#| label: fig-vc-step-response-lelele
#| fig-cap: "Voltage control step response of the designed PLL."
#| fig-align: center

import numpy as np
import matplotlib.pyplot as plt
from scipy.signal import TransferFunction, step

# Numerador e denominador da função de transferência (V_C / ω_R)
num = [(N0 * omega_n_meas**2 / Kvco) * (2 * zeta_meas / omega_n_meas), N0 * omega_n_meas**2 / Kvco]
den = [1, 2 * zeta_meas * omega_n_meas, omega_n_meas**2]

# Criando o sistema (sem multiplicar por s — assume entrada ω_R = Δω/s → degrau em ω_R)
system = TransferFunction(num, den)

# Tempo de simulação
t = np.linspace(0, 300e-6, 1000)

# Resposta ao degrau
t, vc_response = step(system, T=t)
vc_response *= step_m_lin

# Valor final (assumido ou estimado)
final_value = ((f_r+step_m_lin) * N0 - f_o) / Kvco
tol = 0.01 * final_value  # 1% de tolerância
upper_bound = final_value + tol
lower_bound = final_value - tol

# Encontrar settling time: busca o primeiro instante em que o sinal entra e permanece dentro da faixa de tolerância
within_bounds = np.logical_and(vc_response >= lower_bound, vc_response <= upper_bound)

# Janela de amostras para verificar se permanece dentro da tolerância
window_size = 50  # ~5% da simulação

ts_graphical = 0 
for i in range(len(within_bounds) - window_size):
    if np.all(within_bounds[i:i + window_size]):
        ts_graphical = t[i]
        break

# Plotagem
plt.figure(figsize=(8, 4))
plt.plot(t * 1e6, vc_response, label='V_C(t)')
plt.axhline(upper_bound, color='red', linestyle='--', linewidth=0.8, label='+1% Tolerance')
plt.axhline(lower_bound, color='red', linestyle='--', linewidth=0.8, label='-1% Tolerance')

if ts_graphical is not None:
    plt.axvline(ts_graphical * 1e6, color='green', linestyle='--', linewidth=1.2,
                label=f'Settling time: {ts_graphical*1e6:.2f} µs')

plt.title('Resposta à entrada em degrau de frequência Δω (Tensão de controle V_C)')
plt.xlabel('Tempo [µs]')
plt.ylabel('V_C(t) [V]')
plt.grid(True)
plt.legend()
plt.tight_layout()
plt.show()

```

Using the parameters $\omega_n$ and $\zeta$, the transfer function of the entire loop can be described as (neglecting the second capacitor, $C_2$, of the filter):

$$
\frac{\theta_o}{\theta_R}=\frac{\omega_n^2\left(\frac{2\zeta}{\omega_n}s+1\right)}{s^2+2\zeta\omega_n s+\omega_n^2}
$$ {#eq-full-transfer-function-no-C2}

The natural frequency of the loop, $\omega_n$, was computed from the loop bandwidth $\omega_\mathrm{bw}$ and damping factor $\zeta$, yielding a natural frequency $f_n=\omega_n/2\pi=$ `{python} format(f_n)`Hz.

Based on the chosen parameters, the loop filter components were calculated as follows. The primary capacitor of the loop filter is:

$$
C_1 = \frac{K_\theta K_\mathrm{VCO}}{N \omega_n^2}
$$

That is, $C_1=$ `{python} format(C_1)`F.

The second capacitor is chosen to be one-tenth of $C_1$, i.e., $C_2=$ `{python} format(C_2)`F.

The resistor of the loop filter is given by:

$$
R_1 = \frac{\zeta 4 \pi N \omega_n}{I K_\mathrm{VCO}}
$$

That is, $R_1=$ `{python} format(R_1)`$\Omega$.

These component values define the loop filter required to achieve the desired dynamic behavior of the PLL. With these components, it's possible to extract another set of the system parameters, i.e., $f_n=\omega_n/2\pi$, $\zeta$ and $f_\mathrm{bw}=\omega_\mathrm{bw}/2\pi$. The values are $f_n=$ `{python} format(f_n_meas)`Hz, $\zeta=$ `{python} round(zeta_meas, 3)` and $f_\mathrm{bw}=$ `{python} format(f_bw_meas)`Hz.

Similarlly what was presented in [Eq. @eq-full-transfer-function-no-C2], it's possible to present the transfer function of the system in terms of components values. This transfer function can be seen below (this equation uses the second capacitor, $C_2$, of the loop filter, adding a high-frequency pole to the system):

$$
\begin{split}
\frac{\theta_o}{\theta_R}&=\frac{K_\mathrm{VCO}K_\theta\left(1+sC_1R\right)}{s^2N\left(C_1+C_2\right)\left(1+sC_sR\right)+K_\mathrm{VCO}K_\theta\left(1+sC_1R\right)}
\end{split}
$$

where $C_s$ is the paralel between $C_1$ and $C_2$. With the PLL values properly defined, the transfer function of this system is given by:

```{python}
#| echo: false

from scipy import signal
from IPython.display import display, Markdown

def format_eng(x):
    if x == 0:
        return "0"
    exponent = int("{:.0e}".format(x).split('e')[1])
    base = x / (10**exponent)
    return f"{base:.2f}e{exponent}"

def poly_to_str(coeffs):
    terms = []
    degree = len(coeffs) - 1
    for i, coef in enumerate(coeffs):
        power = degree - i
        if abs(coef) < 1e-12:  # ignora coeficientes muito pequenos
            continue
        term = ""
        coef_str = format_eng(coef)
        if not (abs(coef - 1) < 1e-12 and power != 0):
            term += coef_str
        if power > 0:
            term += "s"
            if power > 1:
                term += f"^{power}"
        terms.append(term)
    return " + ".join(terms)

# Em scipy.signal, num e den são arrays dentro de listas (mesmo para SISO)
num = H_sys.num[0][0] if isinstance(H_sys.num[0], (list, tuple)) else H_sys.num
den = H_sys.den[0][0] if isinstance(H_sys.den[0], (list, tuple)) else H_sys.den

num_str = poly_to_str(num)
den_str = poly_to_str(den)

# Exibe como equação no Jupyter
display(Markdown(f"$$H(s) = \\frac{{{num_str}}}{{{den_str}}}$$ {{#eq-transfer-function}}"))

```

```{python}
#| echo: false

from scipy import signal
from IPython.display import display, Markdown

def format_eng(x):
    if x == 0:
        return "0"
    exponent = int("{:.0e}".format(x).split('e')[1])
    base = x / (10**exponent)
    return f"{base:.2f}e{exponent}"

def poly_to_str(coeffs):
    terms = []
    degree = len(coeffs) - 1
    for i, coef in enumerate(coeffs):
        power = degree - i
        if abs(coef) < 1e-12:  # ignora coeficientes muito pequenos
            continue
        term = ""
        coef_str = format_eng(coef)
        if not (abs(coef - 1) < 1e-12 and power != 0):
            term += coef_str
        if power > 0:
            term += "s"
            if power > 1:
                term += f"^{power}"
        terms.append(term)
    return " + ".join(terms)

# Em scipy.signal, num e den são arrays dentro de listas (mesmo para SISO)
num = F_sys.num[0][0] if isinstance(F_sys.num[0], (list, tuple)) else F_sys.num
den = F_sys.den[0][0] if isinstance(F_sys.den[0], (list, tuple)) else F_sys.den

num_str = poly_to_str(num)
den_str = poly_to_str(den)

# Exibe como equação no Jupyter
display(Markdown(f"$$F(s) = \\frac{{{num_str}}}{{{den_str}}}$$ {{#eq-transfer-function-loop-filter}}"))

```

A step with the same amplitude as the previously calculated frequency step was also applied in phase, that is, $\Delta\theta=$ `{python} format_with_pi_latex(step_m_ang_meas)`rad/s. Applying this signal to [Eq. @eq-transfer-function], the response of this system to the phase step is obtained. With this step, the settling time achieved is $t_s=$ `{python} format(t_s_meas)`s. Through simulation, it is possible to verify that the obtained value is $t_s=$ `{python} format(ts_graphical)`s. The simulated response can be visualized in [Fig. @fig-pll-step-response].

```{python}
#| echo: false

import numpy as np
import matplotlib.pyplot as plt
from scipy import signal

# Gerar a resposta em frequência
w, mag, phase = signal.bode(H_sys)

# Encontrar o ponto onde a magnitude cruza -3 dB
cutoff_index = np.argmax(mag <= -3)
cutoff_freq_ang = w[cutoff_index]
cutoff_freq_lin = cutoff_freq_ang/2/np.pi
cutoff_mag = mag[cutoff_index]

# Plotando o diagrama de Bode
plt.figure(figsize=(10, 6))

# Magnitude
plt.subplot(2, 1, 1)
plt.semilogx(w, mag, label='Magnitude (dB)')
plt.axvline(x=cutoff_freq_ang, color='r', linestyle='--', label='Cutoff Frequency (-3 dB)')
plt.axhline(y=-3, color='gray', linestyle=':', label='-3 dB')
plt.scatter(cutoff_freq_ang, cutoff_mag, color='red')  # Marcar o ponto

# Adiciona o valor da frequência no gráfico
plt.text(cutoff_freq_ang, cutoff_mag - 5, f'{format_graph(cutoff_freq_lin)}Hz', color='red',
         ha='center', va='top', fontsize=9, bbox=dict(facecolor='white', edgecolor='red', boxstyle='round'))

plt.title('Frequency Response - Bode Plot')
plt.ylabel('Magnitude (dB)')
plt.grid(True, which="both", ls="--")
plt.legend()

# Fase
plt.subplot(2, 1, 2)
plt.semilogx(w, phase)
plt.axvline(x=cutoff_freq_ang, color='r', linestyle='--')
plt.ylabel('Phase (degrees)')
plt.xlabel('Frequency (rad/s)')
plt.grid(True, which="both", ls="--")

plt.tight_layout()
plt.show()
```

```{python}
#| echo: false

import numpy as np
import matplotlib.pyplot as plt
from scipy import signal

# Gerar a resposta em frequência
w, mag, phase = signal.bode(F_sys)

# Encontrar o ponto onde a magnitude cruza -3 dB
cutoff_index = np.argmax(mag <= -3)
cutoff_freq_ang = w[cutoff_index]
cutoff_freq_lin = cutoff_freq_ang/2/np.pi
cutoff_mag = mag[cutoff_index]

# Plotando o diagrama de Bode
plt.figure(figsize=(10, 6))

# Magnitude
plt.subplot(2, 1, 1)
plt.semilogx(w, mag, label='Magnitude (dB)')

plt.title('Bode Plot - Loop filter')
plt.ylabel('Magnitude (dB)')
plt.grid(True, which="both", ls="--")
plt.legend()

# Fase
plt.subplot(2, 1, 2)
plt.semilogx(w, phase)
plt.ylabel('Phase (degrees)')
plt.xlabel('Frequency (rad/s)')
plt.grid(True, which="both", ls="--")

plt.tight_layout()
plt.show()
```

## Cycle slipping analyzing

```{python}
#| echo: false

import numpy as np

omega_r_safe = omega_n*10
f_r_safe = omega_r_safe/2/np.pi

omega_r_max = omega_n * 2 * np.pi * zeta
f_r_max = omega_r_max/2/np.pi
```

With the intention of conducting tests to verify the occurrence of cycle slipping, the minimum frequency of the reference signal is estimated. A ratio often quoted as being safe is 10:1 for $\omega_r/\omega_n$; thus, the minimum safe frequency for the reference signal is given by $f_r=$ `{python} format(f_r_safe)`Hz. The minimum allowed frequency is given by:

$$
\frac{\omega_r}{\omega_n}\ge 2\pi\zeta
$$

That is, the minimum reference frequency is $f_r\ge$ `{python} format(f_r_max)`Hz. Therefore, since a PLL operating with a reference frequency of $f_r=$ `{python} format(f_r)`Hz and an output frequency of `{python} format(f_o)`Hz is desired, the designed system must operate correctly.

Using the Simulink software, a testbench was prepared to verify the settling time of the PLL. With the previously calculated parameters and using the maximum step frequency value, the resulting curve can be seen in [Fig. @fig-max-step-response]. It can be observed that the step is applied at the instant of 1 ms, as the system is already properly stabilized at that moment. As predicted, cycle slipping does not occur; however, unlike what was calculated, the settling time of this signal is $9.196\;\mu\text{s}$.

<!--![System response to the maximum step frequency](../images/max_step_frequency_response.jpg){#fig-max-step-response}-->

```{python}
#| echo: false

cycle_slip_step_lin = 3*step_m_lin
```

Moreover, contrary to what was previously predicted, for cycle slipping to occur, the step frequency must be approximately $\Delta f=$ `{python} format(cycle_slip_step_lin)`Hz. The system's response to this signal can be seen in [Fig. @fig-cycle-slip-response].

<!--![System response with cycle slip](../images/step_frequency_response_cycle_slip.jpg){#fig-cycle-slip-response}-->

# Behavioral Simulation

# Design Analysis

```{python}
#| echo: false
import math

omega_n=omega_bw/math.sqrt(1+2*zeta**2+math.sqrt(4*zeta**4+4*zeta**2+2))
```

Partindo dos parâmetros já apresentados na introdução, the natural frequency is $\omega_n=$ `{python} format_with_pi_latex(omega_n)`rad/s.

```{python}
#| echo: false
import math

t_m=1/(omega_n*math.sqrt(1-zeta**2))*math.atan(math.sqrt(1-zeta**2)/zeta)
```

Calculando-se o instante em que ocorre a máxima diferença de fase obtém-se, $t_m=$ `{python} format(t_m)`s.

```{python}
#| echo: false
import math

theta_e_m_n = math.sin(omega_n*math.sqrt(1-zeta**2)*t_m)/(math.sqrt(1-zeta**2))*math.e**(-zeta*omega_n*t_m)
theta_e_m = (1.6255) * math.pi
```

And applying $t_m$ to this equation, the maximum normalized phase error in this case is approximately `{python} round(theta_e_m_n, 2)`.

Normalmente, espera-se que o máximo erro de fase suportado pelo PFD é de $2\pi$, entretanto, para coincidir com os valores obtidos com o simulador utilizado, empiricamente obtém-se um maximum phase error, $\theta_{e_m}=$ `{python} format_with_pi_latex(theta_e_m)`.

```{python}
#| echo: false
import math

step_m_ang=theta_e_m*omega_n/theta_e_m_n
step_m_lin = step_m_ang/2/math.pi
```

Substituting the calculated values na [Eq. @eq-maximum-step-frequency] yields $\Delta\omega_m=$ `{python} format_with_pi_latex(step_m_ang)`rad/s or, alternativelly, $\Delta f_m=$ `{python} format(step_m_lin)`Hz. Hence, the maximum input frequency step that still allows theoretical predictability is `{python} format(step_m_lin)`Hz. If the input frequency step exceeds this limit, the PLL may lose lock and exhibit cycle slipping (more details are provided in the next section).

```{python}
#| echo: false
import math

t_s=math.log(0.01*math.sqrt(1-zeta**2), math.e)/(zeta*omega_n)*(-1)
```

Com esse step de frequencia e todos os parâmetros calculados até então, é possível calcular o settling time de 1% de solução, resultando em $t_s=$ `{python} format(t_s)`s.

## Project

```{python}
#| echo: false

import numpy as np

# Parâmetros iniciais
#f_r = 10e6 # Hz
#omega_r = 2 * np.pi * f_r # rad/s
#f_o = 50e6 # Hz
#omega_o = 2 * np.pi * f_r # rad/s
#N = 5
#Kvco = KVCO = 10e6 # Hz/V
#Kvco_ang = 2 * np.pi * Kvco # rad/s/V
#I = 1e-3
#K_theta = Kphase = I / (2 * np.pi) # A/rad
#A0 = 6.258
#K = A0 * Kphase * KVCO / N
```

Para exemplificar os resultados obtidos até então, an integer-N Phase-Locked Loop (PLL) was developed with a division factor of $N=$ `{python} N`. The system was designed with a VCO gain of $K_\mathrm{VCO}=$ `{python} format(KVCO)`Hz/V, and a charge pump current of $I=$ `{python} format(I)`A. The phase detector gain is given by $K_\theta=I/2\pi$. Além disso, para que seja possível coincidir os valores calculados com as respostas obtidas na simulação, foi obtido um valor empírico para o ganho dc, $A_0=$ `{python} round(A0, 2)`.

```{python}
#| echo: false

from scipy import signal

omega_n = omega_bw / np.sqrt(1 + 2 * zeta**2 + np.sqrt(4 * zeta**4 + 4 * zeta**2 + 2))
f_n = omega_n/2/np.pi
```

The natural frequency of the loop, $\omega_n$, was computed from the loop bandwidth $\omega_\mathrm{bw}$ and damping factor $\zeta$, yielding a natural frequency $f_n=\omega_n/2\pi=$ `{python} format(f_n)`Hz.

```{python}
#| echo: false

# Parâmetros do PLL
C1 = K / omega_n**2
C2 = C1 / 10
R1 = 2 * zeta * omega_n / K
```

Based on the chosen parameters, the loop filter components were calculated as follows. The primary capacitor of the loop filter is given by the [Eq. @], that is, $C_1=$ `{python} format(C1)`F. The second capacitor is chosen to be one-tenth of $C_1$, i.e., $C_2=$ `{python} format(C2)`F. The resistor of the loop filter is given by [Eq. @], that is, $R_1=$ `{python} format(R1)`$\Omega$.

```{python}
#| echo: false

import math
import numpy as np

# Análise de resposta em frequência
omega_n_meas = np.sqrt(K / C1)
f_n_meas = omega_n_meas / (2 * np.pi)

zeta_meas = (R1 / 2) * np.sqrt(K * C1)

t_m_meas = 1 / (omega_n_meas * math.sqrt(1 - zeta_meas**2)) * math.atan(math.sqrt(1 - zeta_meas**2) / zeta_meas)

theta_e_m_n_meas = math.sin(omega_n_meas * math.sqrt(1 - zeta_meas**2) * t_m_meas) / \
                   (math.sqrt(1 - zeta_meas**2)) * math.e**(-zeta_meas * omega_n_meas * t_m_meas)

theta_e_m_meas = theta_e_m

omega_bw_meas = omega_n_meas * np.sqrt(1 + 2 * zeta_meas**2 + np.sqrt(4 * zeta_meas**4 + 4 * zeta_meas**2 + 2))
f_bw_meas = omega_bw_meas / (2 * np.pi)

step_m_ang_meas = theta_e_m_meas * omega_n_meas / theta_e_m_n_meas
step_m_lin_meas = step_m_ang_meas / (2 * math.pi)

t_s_meas = math.log(0.01 * math.sqrt(1 - zeta_meas**2)) / (zeta_meas * omega_n_meas) * (-1)

```

These component values define the loop filter required to achieve the desired dynamic behavior of the PLL. With these components, it's possible to extract another set of the system parameters, i.e., $f_n=\omega_n/2\pi$, $\zeta$ and $f_\mathrm{bw}=\omega_\mathrm{bw}/2\pi$. The values are $f_n=$ `{python} format(f_n_meas)`Hz, $\zeta=$ `{python} round(zeta_meas, 3)` and $f_\mathrm{bw}=$ `{python} format(f_bw_meas)`Hz.

```{python}
#| echo: false

from scipy import signal
from IPython.display import display, Markdown

def format_eng(x):
    if x == 0:
        return "0"
    exponent = int("{:.0e}".format(x).split('e')[1])
    base = x / (10**exponent)
    return f"{base:.2f}e{exponent}"

def poly_to_str(coeffs):
    terms = []
    degree = len(coeffs) - 1
    for i, coef in enumerate(coeffs):
        power = degree - i
        if abs(coef) < 1e-12:  # ignora coeficientes muito pequenos
            continue
        term = ""
        coef_str = format_eng(coef)
        if not (abs(coef - 1) < 1e-12 and power != 0):
            term += coef_str
        if power > 0:
            term += "s"
            if power > 1:
                term += f"^{power}"
        terms.append(term)
    return " + ".join(terms)
```

A transfer function of the filter is given by:

```{python}
#| echo: false

# Numerador e denominador do compensador F(s)
# F(s) = (s*C1*R1 + 1) / (s^2*C1*C2*R1 + s*(C1 + C2))
F_num = [C1 * R1, 1]
F_den = [C1 * C2 * R1, C1 + C2, 0]

F_sys = signal.TransferFunction(F_num, F_den)
```

```{python}
#| echo: false
# Em scipy.signal, num e den são arrays dentro de listas (mesmo para SISO)
num = F_sys.num[0][0] if isinstance(F_sys.num[0], (list, tuple)) else F_sys.num
den = F_sys.den[0][0] if isinstance(F_sys.den[0], (list, tuple)) else F_sys.den

num_str = poly_to_str(num)
den_str = poly_to_str(den)

# Exibe como equação no Jupyter
display(Markdown(f"$$F(s) = \\frac{{{num_str}}}{{{den_str}}}$$ {{#eq-filter-loop-transfer-function}}"))

```

The open-loop transfer function of the system is given by:

```{python}
#| echo: false
# F(s) * K / s = G(s)
G_num = np.polymul(F_num, [K])
G_den = np.polymul(F_den, [1, 0])  # Multiplica por s

G_sys = signal.TransferFunction(G_num, G_den)
```

```{python}
#| echo: false
# Em scipy.signal, num e den são arrays dentro de listas (mesmo para SISO)
num = G_sys.num[0][0] if isinstance(G_sys.num[0], (list, tuple)) else G_sys.num
den = G_sys.den[0][0] if isinstance(G_sys.den[0], (list, tuple)) else G_sys.den

num_str = poly_to_str(num)
den_str = poly_to_str(den)

# Exibe como equação no Jupyter
display(Markdown(f"$$G(s) = \\frac{{{num_str}}}{{{den_str}}}$$ {{#eq-open-loop-num-transfer-function}}"))
```

The closed-loop transfer function of the entire system, both in frequency domain and phase domain is given by:

```{python}
#| echo: false
# H(s) = G(s) / (1 + G(s))
H_num = G_num
H_den = np.polyadd(G_den, G_num)

# Criando o sistema de tempo contínuo
H_sys = signal.TransferFunction(H_num, H_den)
```

```{python}
#| echo: false
# Em scipy.signal, num e den são arrays dentro de listas (mesmo para SISO)
num = H_sys.num[0][0] if isinstance(H_sys.num[0], (list, tuple)) else H_sys.num
den = H_sys.den[0][0] if isinstance(H_sys.den[0], (list, tuple)) else H_sys.den

num_str = poly_to_str(num)
den_str = poly_to_str(den)

# Exibe como equação no Jupyter
display(Markdown(f"$$H(s) = \\frac{{{num_str}}}{{{den_str}}}$$ {{#eq-closed-loop-num-transfer-function}}"))
```

The transfer function of the output phase is given by:

```{python}
#| echo: false

# Agora temos o denominador completo
# O_sys = H_sys * N
# Para isso, O_sys = (H_num * N) / (H_den)
num_O = np.convolve(H_sys.num, [N0])
den_O = np.convolve(H_sys.den, [1])

# Criar a função de transferência final
O_sys = signal.TransferFunction(num_O, den_O)
```

```{python}
#| echo: false
# Em scipy.signal, num e den são arrays dentro de listas (mesmo para SISO)
num = O_sys.num[0][0] if isinstance(O_sys.num[0], (list, tuple)) else O_sys.num
den = O_sys.den[0][0] if isinstance(O_sys.den[0], (list, tuple)) else O_sys.den

num_str = poly_to_str(num)
den_str = poly_to_str(den)

# Exibe como equação no Jupyter
display(Markdown(f"$$O(s) = \\frac{{{num_str}}}{{{den_str}}}$$ {{#eq-output-phase-num-transfer-function}}"))
```

The transfer function of the phase error is given by:

```{python}
#| echo: false

# Kvco/s => [Kvco] / [1, 0]
Kvco_over_s = ([Kvco], [1, 0])

# F_sys => ([1], [1, 1]) já está
# Multiplicando numeradores e denominadores via convolução
num_den = np.convolve([1/N0], Kvco_over_s[0])
num_den = np.convolve(num_den, F_sys.num)
num_den = np.convolve(num_den, [K_theta])
num_den = np.convolve(num_den, [A0])

den_den = np.convolve(Kvco_over_s[1], F_sys.den)

# Agora temos o denominador completo
# E_sys = H_sys / (den)
# Para isso, E_sys = (H_num * den_den) / (H_den * den_num)
num_E = np.convolve(H_sys.num, den_den)
den_E = np.convolve(H_sys.den, num_den)

# Criar a função de transferência final
E_sys = signal.TransferFunction(num_E, den_E)
```

```{python}
#| echo: false
# Em scipy.signal, num e den são arrays dentro de listas (mesmo para SISO)
num = E_sys.num[0][0] if isinstance(E_sys.num[0], (list, tuple)) else E_sys.num
den = E_sys.den[0][0] if isinstance(E_sys.den[0], (list, tuple)) else E_sys.den

num_str = poly_to_str(num)
den_str = poly_to_str(den)

# Exibe como equação no Jupyter
display(Markdown(f"$$E(s) = \\frac{{{num_str}}}{{{den_str}}}$$ {{#eq-phase-error-num-transfer-function}}"))
```

The transfer function of the voltage control is given by:

```{python}
#| echo: false

# Vc_sys = H_sys / ((1/N)*(Kvco/s))
#Kvco_over_s = ([Kvco], [1, 0])
Kvco_over_s = ([Kvco], [1])
den_num = np.convolve([1/N0], Kvco_over_s[0])
den_den = Kvco_over_s[1]

num_Vc = np.convolve(H_sys.num, den_den)
den_Vc = np.convolve(H_sys.den, den_num)

Vc_sys = signal.TransferFunction(num_Vc, den_Vc)
```

```{python}
#| echo: false
# Em scipy.signal, num e den são arrays dentro de listas (mesmo para SISO)
num = Vc_sys.num[0][0] if isinstance(Vc_sys.num[0], (list, tuple)) else Vc_sys.num
den = Vc_sys.den[0][0] if isinstance(Vc_sys.den[0], (list, tuple)) else Vc_sys.den

num_str = poly_to_str(num)
den_str = poly_to_str(den)

# Exibe como equação no Jupyter
display(Markdown(f"$$V_C(s) = \\frac{{{num_str}}}{{{den_str}}}$$ {{#eq-voltage-control-num-transfer-function}}"))
```

## Validade do modelo

```{python}
#| echo: false

import numpy as np

omega_r_safe = omega_n*10
f_r_safe = omega_r_safe/2/np.pi

omega_r_max = omega_n * 2 * np.pi * zeta
f_r_max = omega_r_max/2/np.pi
```

The minimum safe frequency for the reference signal is given by $f_r=$ `{python} format(f_r_safe)`Hz. However, the minimum reference frequency is $f_r\ge$ `{python} format(f_r_max)`Hz. Therefore, since a PLL operating with a reference frequency of $f_r=$ `{python} format(f_r)`Hz and an output frequency of `{python} format(f_o)`Hz is desired, the designed system must operate correctly.

# System Responses

```{python}
#| echo: false
#| label: fig-pll-step-response
#| fig-cap: "Step Response of the Closed-Loop Frequency Transfer Function of the PLL."
#| fig-align: center

import numpy as np
import matplotlib.pyplot as plt
from scipy import signal

# Tempo de simulação
t = np.linspace(0, 300e-6, 100000)

# Criar vetor da entrada (degrau de amplitude 5)
step_amplitude = 0.5e6
u = np.ones_like(t) * step_amplitude

# Simulação da resposta ao degrau via lsim
t_out, y_out, _ = signal.lsim(H_sys, U=u, T=t)

# Valor final da resposta
final_value = y_out[-1]
tol = 0.01 * final_value  # 1% da resposta final

# Encontrar o tempo em que a resposta entra e permanece na faixa de ±1%
ts_graphical = 0
within_bounds = np.abs(y_out - u) <= tol
for i in range(len(within_bounds)):
    if all(within_bounds[i:]):
        ts_graphical = t_out[i]
        break

# Gerar gráfico
plt.figure(figsize=(8, 5))
plt.plot(t_out * 1e6, y_out, label='Step Response')
plt.plot(t_out * 1e6, u, linestyle=':', color='black', label='Step')
plt.axhline(final_value + tol, color='red', linestyle='--', linewidth=0.8, label='±1% Tolerance')
plt.axhline(final_value - tol, color='red', linestyle='--', linewidth=0.8)
plt.axvline(ts_graphical * 1e6, color='green', linestyle='--', linewidth=1.2, label=f'Graphical settling time: {ts_graphical*1e6:.5f} us')
plt.xlabel('Time [us]')
plt.ylabel('Frequency (Hz)')
plt.grid(True)
plt.legend()
plt.tight_layout()
plt.show()
```

A step with the same amplitude as the previously calculated frequency step was also applied in phase, that is, $\Delta\theta=$ `{python} format_with_pi_latex(step_m_ang_meas)`rad/s. Applying this signal to [Eq. @eq-transfer-function], the response of this system to the phase step is obtained. With this step, the settling time achieved is $t_s=$ `{python} format(t_s_meas)`s. Through simulation, it is possible to verify that the obtained value is $t_s=$ `{python} format(ts_graphical)`s. The simulated response can be visualized in [Fig. @fig-pll-step-response].

```{python}
#| echo: false
#| label: fig-pll-step-response-alt
#| fig-cap: "Closed-loop phase step response of the designed PLL."
#| fig-align: center

import numpy as np
import matplotlib.pyplot as plt
from scipy import signal

# Tempo de simulação
t = np.linspace(0, 500e-6, 100000)

# Criar vetor da entrada (degrau de amplitude 5)
step_amplitude = omega_r * t
u = np.ones_like(t) * step_amplitude

# Simulação da resposta ao degrau via lsim
t_out, y_out, _ = signal.lsim(H_sys, U=u, T=t)

# Valor final da resposta
final_value = y_out[-1]
tol = 0.01 * final_value  # 1% da resposta final

# Encontrar o tempo em que a resposta entra e permanece na faixa de ±1%
ts_graphical = 0
within_bounds = np.abs(y_out - u) <= tol
for i in range(len(within_bounds)):
    if all(within_bounds[i:]):
        ts_graphical = t_out[i]
        break

# Gerar gráfico
plt.figure(figsize=(8, 5))
plt.plot(t_out * 1e6, y_out, label='Ramp Response')
plt.plot(t_out * 1e6, u, linestyle=':', color='black', label='Step')
plt.axvline(ts_graphical * 1e6, color='green', linestyle='--', linewidth=1.2, label=f'Graphical settling time: {ts_graphical*1e6:.5f} us')
plt.xlabel('Time [us]')
plt.ylabel('Phase [rad]')
plt.title('Step Response with Settling Time (99%)')
plt.grid(True)
plt.legend()
plt.tight_layout()
plt.show()
```

```{python}
#| echo: false
#| label: fig-pll-step-iasdhf-alt
#| fig-cap: "Closed-loop kjas step response of the designed PLL."
#| fig-align: center

import numpy as np
import matplotlib.pyplot as plt
from scipy import signal

# Tempo de simulação
t = np.linspace(0, 500e-6, 100000)

# Criar vetor da entrada (degrau de amplitude 5)
step_amplitude = omega_r * t
u = np.ones_like(t) * step_amplitude

# Simulação da resposta ao degrau via lsim
t_out, y_out, _ = signal.lsim(H_sys, U=u, T=t)
t_out, y_out2, _ = signal.lsim(O_sys, U=u, T=t)

# Gerar gráfico
plt.figure(figsize=(8, 5))
plt.plot(t_out * 1e6, y_out, label='Ramp Response')
plt.plot(t_out * 1e6, y_out2, label='Ramp N Response')
plt.plot(t_out * 1e6, u, linestyle=':', color='black', label='Step')
plt.xlabel('Time [us]')
plt.ylabel('Phase [rad]')
plt.title('Step Response with Settling Time (99%)')
plt.grid(True)
plt.legend()
plt.tight_layout()
plt.show()
```

```{python}
#| echo: false
#| label: fig-pll-step-response-alt-double-ramp
#| fig-cap: "Closed-loop phase step response of the designed PLL."
#| fig-align: center

import numpy as np
import matplotlib.pyplot as plt
from scipy import signal

# Tempo de simulação
t = np.linspace(0, 500e-6, 100000)

# Frequências
f_base = omega_r  # 10 MHz
f_step = step_m_ang_meas  # step de 10 MHz
t_step = 20e-6  # instante do step

# Entrada em fase: integral da frequência
phi_in = np.where(
    t < t_step,
    f_base * t,
    f_base * t + f_step * (t - t_step)
)

# Simulação da resposta
t_out, y_out, _ = signal.lsim(H_sys, U=phi_in, T=t)

### --- 1º Settling Time (antes do step) ---
# Índices até o step
idx_step = np.searchsorted(t_out, t_step)
y_before = y_out[:idx_step]
u_before = phi_in[:idx_step]

# Valor final antes do step
final_val_1 = y_before[-1]
tol_1 = 0.01 * final_val_1

# Encontrar settling time 1
settling_instant1 = 0
within_1 = np.abs(y_before - u_before) <= tol_1
for i in range(len(within_1)):
    if all(within_1[i:]):
        settling_instant1 = t_out[i]
        break

### --- 2º Settling Time (após o step) ---
y_after = y_out[idx_step:]
u_after = phi_in[idx_step:]
t_after = t_out[idx_step:]

# Valor final após o step
final_val_2 = y_after[-1]
tol_2 = 0.01 * final_val_2

# Encontrar settling time 2
settling_instant2 = 0
within_2 = np.abs(y_after - u_after) <= tol_2
for i in range(len(within_2)):
    if all(within_2[i:]):
        settling_instant2 = t_after[i]
        break

settling_time1 = settling_instant1
settling_time2 = settling_instant2 - t_step

### --- Gráfico ---
plt.figure(figsize=(8, 5))
plt.plot(t_out * 1e6, y_out, label='Output Phase')
plt.plot(t_out * 1e6, phi_in, linestyle=':', color='black', label='Input Phase')

plt.axvline(settling_instant1 * 1e6, color='green', linestyle='--', linewidth=1.2, label=f'Settling time 1: {settling_time1*1e6:.2f} µs')
plt.axvline(settling_instant2 * 1e6, color='blue', linestyle='--', linewidth=1.2, label=f'Settling time 2: {settling_time2*1e6:.2f} µs')
plt.axvline(t_step * 1e6, color='purple', linestyle='--', linewidth=1.0, label='Step at 20 µs')

plt.xlabel('Time [µs]')
plt.ylabel('Phase [rad]')
plt.title('PLL Response with Two Settling Times')
plt.grid(True)
plt.legend()
plt.tight_layout()
plt.show()

```

```{python}
#| echo: false
#| label: fig-pll-step-response-1
#| fig-cap: "Closed-loop step response of the designed PLL."
#| fig-align: center

import numpy as np
import matplotlib.pyplot as plt
from scipy import signal

# Tempo de simulação
t = np.linspace(0, 500e-6, 100000)

# Entrada: rampa em fase (step de 10 MHz em frequência)
# Fase = ∫ frequência ⇒ rampa
freq_step = omega_r  # Step de 10 MHz em frequência
phi_in = freq_step * t  # Rampa de fase

# Simulação da resposta do sistema à rampa
t_out, phi_out, _ = signal.lsim(H_sys, U=phi_in, T=t)

# Erro de seguimento
tracking_error = phi_in - phi_out

# Definir tolerância de erro (ex: 1% da entrada)
tol = 0.01 * np.max(phi_in)

# Encontrar tempo de seguimento gráfico
tracking_ts = 0
within_bounds = np.abs(tracking_error) <= tol
for i in range(len(within_bounds)):
    if all(within_bounds[i:]):
        tracking_ts = t_out[i]
        break

# Plotagem
plt.figure(figsize=(10, 6))
plt.plot(t_out * 1e6, phi_in, 'k--', label='Fase de entrada (rampa)')
plt.plot(t_out * 1e6, phi_out, 'b', label='Fase de saída')
plt.plot(t_out * 1e6, tracking_error, 'r', label='Erro de seguimento')
plt.axhline(tol, color='gray', linestyle='--', linewidth=0.8, label='±1% Tolerância')
plt.axhline(-tol, color='gray', linestyle='--', linewidth=0.8)
plt.axvline(tracking_ts * 1e6, color='green', linestyle='--', linewidth=1.2,
            label=f'Tempo de seguimento: {tracking_ts*1e6:.2f} us')
plt.title('Resposta do PLL a uma rampa de fase (step em frequência)')
plt.xlabel('Tempo [us]')
plt.ylabel('Fase [rad]')
plt.grid(True)
plt.legend()
plt.tight_layout()
plt.show()

```

```{python}
#| echo: false
#| label: fig-phase-error-step-response
#| fig-cap: "Phase error response of step frequency."
#| fig-align: center

import numpy as np
import matplotlib.pyplot as plt
from scipy import signal

# Tempo de simulação
t = np.linspace(0, 300e-6, 100000)

# Criar vetor da entrada (degrau de amplitude 5)
step_amplitude = step_m_ang_meas * t
u = np.ones_like(t) * step_amplitude

# Simulação da resposta ao degrau via lsim
t_out, y_out, _ = signal.lsim(E_sys, U=u, T=t)

# Cálculo do valor máximo (pico)
max_value = np.max(np.abs(y_out))
tol = 0.01 * max_value  # 1% do valor de pico (centrado em zero)

# Novo cálculo do tempo de acomodação (settling time)
within_bounds = np.abs(y_out) <= tol
ts_graphical_peak = 0
for i in range(len(within_bounds)):
    if all(within_bounds[i:]):
        ts_graphical_peak = t_out[i]
        break

# Gerar gráfico
plt.figure(figsize=(8, 5))
plt.plot(t_out * 1e6, y_out, label='Step Response')

# Linha no valor máximo (positivo)
plt.axhline(max_value, color='blue', linestyle=':', linewidth=1.2, label=f'Max Value: {max_value:.3f}')

# Linhas horizontais em ±1% do valor de pico (em torno de zero)
plt.axhline(tol, color='purple', linestyle='--', linewidth=1.0, label=f'+1% Peak ({tol:.3e})')
plt.axhline(-tol, color='purple', linestyle='--', linewidth=1.0, label=f'-1% Peak ({-tol:.3e})')

# Marcação do novo tempo de acomodação
plt.axvline(ts_graphical_peak * 1e6, color='green', linestyle='--', linewidth=1.2, label=f'Settling Time (±1% Peak): {ts_graphical_peak*1e6:.3f} μs')

plt.xlabel('Time [us]')
plt.ylabel('Phase Error [rad]')
plt.title('Step Response with Settling Time (±1% of Peak)')
plt.grid(True)
plt.legend()
plt.tight_layout()
plt.show()

```

```{python}
#| echo: false
#| label: fig-pll-step-response-2
#| fig-cap: "Closed-loop step response of the designed PLL."
#| fig-align: center

import numpy as np
import matplotlib.pyplot as plt
from scipy import signal

# Tempo de simulação
t = np.linspace(0, 8/omega_n_meas, 100000)

# Criar vetor da entrada (degrau de amplitude 5)
step_amplitude = t
u = np.ones_like(t) * step_amplitude

# Simulação da resposta ao degrau via lsim
t_out, y_out, _ = signal.lsim(E_sys, U=u, T=t)

y_out *= omega_n_meas

# Cálculo do valor máximo (pico)
max_value = np.max(np.abs(y_out))
tol = 0

# Gerar gráfico
plt.figure(figsize=(8, 5))
plt.plot(t_out * omega_n_meas, y_out, label='Step Response')

# Linha no valor máximo (positivo)
plt.axhline(max_value, color='blue', linestyle=':', linewidth=1.2, label=f'Max Value: {max_value:.3f}')

# Linhas horizontais em ±1% do valor de pico (em torno de zero)
plt.axhline(tol, color='purple', linestyle='--', linewidth=1.0, label=f'+1% Peak ({tol:.3e})')

plt.xlabel('Time [us]')
plt.ylabel('Phase Error [rad]')
plt.title('Step Response with Settling Time (±1% of Peak)')
plt.grid(True)
plt.legend()
plt.tight_layout()
plt.show()

```

```{python}
#| echo: false
#| label: fig-pll-step-response-3
#| fig-cap: "Closed-loop step response of the designed PLL."
#| fig-align: center

from scipy import signal
import numpy as np

# Supondo que você já tem H_sys
# H_sys = signal.TransferFunction(num, den)

# 1. Obtenha os polos do sistema
poles = H_sys.poles

# 2. Pegue o par de polos complexos conjugados dominante
#    (normalmente com parte imaginária não nula)
complex_poles = [p for p in poles if np.imag(p) != 0]
if not complex_poles:
    raise ValueError("Não há polos complexos — sistema não oscilatório ou não de 2ª ordem.")

# 3. Escolha um dos polos complexos
p = complex_poles[0]

# 4. Calcule a frequência natural (ωn) e o fator de amortecimento (ζ)
omega_n = np.abs(p)
zeta = -np.real(p) / omega_n

# 5. Imprima os resultados
#print(f"Polo dominante: {p}")
#print(f"Frequência natural ωn: {omega_n:.2e} rad/s")
print(f"Constante de amortecimento ζ: {zeta:.3f}")

```

```{python}
#| echo: false
#| label: fig-vc-step-response
#| fig-cap: "Voltage control step response of the designed PLL."
#| fig-align: center

import numpy as np
import matplotlib.pyplot as plt
from scipy.signal import TransferFunction, step

# Tempo de simulação
t = np.linspace(0, 300e-6, 1000)

# Resposta ao degrau
t, vc_response = step(Vc_sys, T=t)
vc_response *= step_m_lin_meas

# Valor final (assumido ou estimado)
final_value = ((f_r+step_m_lin_meas) * N0 - f_o) / KVCO
#final_value = 0.5
tol = 0.01 * final_value  # 1% de tolerância
upper_bound = final_value + tol
lower_bound = final_value - tol

# Encontrar settling time: busca o primeiro instante em que o sinal entra e permanece dentro da faixa de tolerância
within_bounds = np.logical_and(vc_response >= lower_bound, vc_response <= upper_bound)

# Janela de amostras para verificar se permanece dentro da tolerância
window_size = 150  # ~5% da simulação

ts_graphical = 0
for i in range(len(within_bounds) - window_size):
    if np.all(within_bounds[i:i + window_size]):
        ts_graphical = t[i]
        break

# Plotagem
plt.figure(figsize=(8, 4))
plt.plot(t * 1e6, vc_response, label='V_C(t)')
plt.axhline(upper_bound, color='red', linestyle='--', linewidth=0.8, label='+1% Tolerance')
plt.axhline(lower_bound, color='red', linestyle='--', linewidth=0.8, label='-1% Tolerance')

if ts_graphical is not None:
    plt.axvline(ts_graphical * 1e6, color='green', linestyle='--', linewidth=1.2,
                label=f'Settling time: {ts_graphical*1e6:.2f} µs')

plt.title('Resposta à entrada em degrau de frequência Δω (Tensão de controle V_C)')
plt.xlabel('Tempo [µs]')
plt.ylabel('V_C(t) [V]')
plt.grid(True)
plt.legend()
plt.tight_layout()
plt.show()

```

```{python}
#| echo: false
#| label: fig-pll-step-response-4
#| fig-cap: "Closed-loop step response of the designed PLL."
#| fig-align: center

import numpy as np
import matplotlib.pyplot as plt
from scipy.signal import TransferFunction, step

# Tempo de simulação
t = np.linspace(0, 8/omega_n_meas, 1000)

# Resposta ao degrau
t, vc_response = step(Vc_sys, T=t)
vc_response *= Kvco / N

# Valor final (assumido ou estimado)
final_value = ((f_r+step_m_lin) * N0 - f_o) / KVCO
final_value *= KVCO / (step_m_lin_meas * N0)
#final_value = 0.5

# Janela de amostras para verificar se permanece dentro da tolerância
window_size = 50  # ~5% da simulação

# Plotagem
plt.figure(figsize=(8, 4))
plt.plot(t * omega_n_meas, vc_response, label='V_C(t)')
plt.axhline(final_value, color='red', linestyle='--', linewidth=0.8, label='Valor desejado')

plt.title('Resposta à entrada em degrau de frequência Δω (Tensão de controle V_C)')
plt.xlabel('Tempo [µs]')
plt.ylabel('V_C(t) [V]')
plt.grid(True)
plt.legend()
plt.tight_layout()
plt.show()

```

```{python}
#| echo: false
#| label: fig-pll-step-response-5
#| fig-cap: "Closed-loop step response of the designed PLL."
#| fig-align: center

import numpy as np
import matplotlib.pyplot as plt
from scipy import signal

# Gerar a resposta em frequência
w, mag, phase = signal.bode(H_sys)

# Encontrar o ponto onde a magnitude cruza -3 dB
cutoff_index = np.argmax(mag <= -3)
cutoff_freq_ang = w[cutoff_index]
cutoff_freq_lin = cutoff_freq_ang/2/np.pi
cutoff_mag = mag[cutoff_index]

# Plotando o diagrama de Bode
plt.figure(figsize=(10, 6))

# Magnitude
plt.subplot(2, 1, 1)
plt.semilogx(w, mag, label='Magnitude (dB)')
plt.axvline(x=cutoff_freq_ang, color='r', linestyle='--', label='Cutoff Frequency (-3 dB)')
plt.axhline(y=-3, color='gray', linestyle=':', label='-3 dB')
plt.scatter(cutoff_freq_ang, cutoff_mag, color='red')  # Marcar o ponto

# Adiciona o valor da frequência no gráfico
plt.text(cutoff_freq_ang, cutoff_mag - 5, f'{format_graph(cutoff_freq_lin)}Hz', color='red',
         ha='center', va='top', fontsize=9, bbox=dict(facecolor='white', edgecolor='red', boxstyle='round'))

plt.title('Frequency Response - Bode Plot')
plt.ylabel('Magnitude (dB)')
plt.grid(True, which="both", ls="--")
plt.legend()

# Fase
plt.subplot(2, 1, 2)
plt.semilogx(w, phase)
plt.axvline(x=cutoff_freq_ang, color='r', linestyle='--')
plt.ylabel('Phase (degrees)')
plt.xlabel('Frequency (rad/s)')
plt.grid(True, which="both", ls="--")

plt.tight_layout()
plt.show()
```

```{python}
#| echo: false
#| label: fig-pll-step-response-6
#| fig-cap: "Closed-loop step response of the designed PLL."
#| fig-align: center

import numpy as np
import matplotlib.pyplot as plt
from scipy import signal

# Gerar a resposta em frequência
w, mag, phase = signal.bode(F_sys)

# Encontrar o ponto onde a magnitude cruza -3 dB
cutoff_index = np.argmax(mag <= -3)
cutoff_freq_ang = w[cutoff_index]
cutoff_freq_lin = cutoff_freq_ang/2/np.pi
cutoff_mag = mag[cutoff_index]

# Plotando o diagrama de Bode
plt.figure(figsize=(10, 6))

# Magnitude
plt.subplot(2, 1, 1)
plt.semilogx(w, mag, label='Magnitude (dB)')

plt.title('Bode Plot - Loop filter')
plt.ylabel('Magnitude (dB)')
plt.grid(True, which="both", ls="--")
plt.legend()

# Fase
plt.subplot(2, 1, 2)
plt.semilogx(w, phase)
plt.ylabel('Phase (degrees)')
plt.xlabel('Frequency (rad/s)')
plt.grid(True, which="both", ls="--")

plt.tight_layout()
plt.show()
```

# Conclusion

This report has outlined the key steps in designing an integer-N frequency synthesizer operating within the specified frequency range. A continuous-time model was adopted due to the large separation between the reference frequency and the loop bandwidth. Component values for the VCO and loop filter were estimated based on reasonable assumptions, providing a solid foundation for subsequent implementation and simulation.

# References {-}