### 2025-06-02 CaseStudies/Example3_2_LimitsoftheTheorySoFar/Report/article/article.qmd
Finishing articles .

---
title: Exploring the Validity Range of PLL Analytical Models
subtitle: Case Study - Example 3.2 from the Reference Material
author:
  - name: Bruno Alexandre Fraga
    email: bruno.fraga@posgrad.ufsc.br
    affiliations: 
        - id: some-tech
          name: Universidade Federal de Santa Catarina
          department: Departamento de Engenharia Elétrica e Eletrônica
          address: R. Delfino Conti
          city: Florianópolis
          state: Santa Catarina
          postal-code: 88040-370
abstract: |
    This report presents the analytical development and performance estimation of a second-order Integer-N Phase-Locked Loop (PLL) based on a Phase-Frequency Detector (PFD) and a Charge Pump (CP). The loop filter is designed to meet specified damping and bandwidth characteristics. The analysis determines the maximum allowable frequency step at the input that ensures the PLL remains in lock, based on the system's maximum normalized phase error. Additionally, the settling time following such a frequency step is estimated using classical control theory. The results provide useful insights into the dynamic behavior of the loop, serving as a theoretical benchmark for future simulation and implementation efforts.
keywords: 
    - Integer-N PLL
    - Loop Dynamics
    - Frequency Step
    - Settling Time 
date: last-modified
bibliography:
  - bibliography.bib
  - zoterobibliography.bib
format:
  elsevier-pdf:
    include-in-header:
      - text: |
          \usepackage{steinmetz}
          \usepackage[a4paper, margin=1in]{geometry}
          \usepackage{graphicx} % Required for inserting images
          \usepackage{circuitikz}
          \usepackage{amsmath}
          \usepackage{amssymb}
          \usepackage{amsthm}
          \usepackage{comment}
          \usepackage{tikz}
          \usetikzlibrary{shapes,arrows}
          \usetikzlibrary{shapes,arrows,positioning,calc}

          \newcommand{\parallelsum}{\mathbin{\!/\mkern-5mu/\!}}
    keep-tex: true
    journal:
      name: Programa de Pós-Graduação em Engenharia Elétrica
      formatting: preprint
      # model: 3p # Don't set a model with preprint
      cite-style: number
---

```{python}
#| echo: false

from format_utils import format, format_graph, format_no_markdown, format_with_pi_latex

```

# Introduction

```{python}
#| echo: false
import math

zeta = 0.707
f_bw = 150e3
omega_bw = 2*math.pi*f_bw
```

This report presents the development of an Integer-N Phase-Frequency Detector/Charge Pump-based Phase-Locked Loop (PFD/CP-based PLL) with a second-order loop filter. The loop filter is designed so that the loop bandwidth has a damping factor $\zeta=$ `{python} zeta`, and a 3-dB bandwidth of $f_{3\text{dB}}=$ `{python} format(f_bw)`Hz. The example analyzed in this report is based on the case presented in @rogers2006IntegratedCircuitDesign.

## Objective

The objective of this study is to answer the following two questions:

- What is the maximum frequency step at the input such that the current theoretical model remains valid in predicting the system's behavior?
- Assuming this condition is satisfied, how long will the system take to settle following a frequency step?

# System Modeling

The [Fig. @fig-PLL-integerN-blocks] presents the block diagram of an Integer-N PFD/CP-based PLL.

![Block diagram of Integer-N](../images/PLL_integerN_blocks.png){#fig-PLL-integerN-blocks}

The second-order loop filter transfer function, $F(s)$ can be described as:

$$
F(s) = \left(R+\frac{1}{sC_1}\right)\parallelsum\frac{1}{sC_2}
$$ {#eq-tf-filter}

The open-loop transfer function of the system shown in \[Fig. @fig-PLL-integerN-blocks] can be written as follows:

$$
G(s) = \frac{KF(s)}{s}
$$ {#eq-tf-open-loop}

where

$$
K = A_0K_\theta K_\mathrm{VCO}N^{-1}
$$

such that $A_0$ is the DC gain (usually unity), $K_\theta = I / 2\pi$ is the charge pump gain, $K_\mathrm{VCO}$ is the VCO gain, and $N$ is the frequency divider. From there, it is possible to obtain the closed-loop transfer function of the system, given by:

$$
H(s)=\frac{\theta_\mathrm{fb}}{\theta_R}=\frac{\omega_\mathrm{fb}}{\omega_R}=\frac{G(s)}{1+G(s)}
$$ {#eq-tf-closed-loop}

It can be observed that the transfer function $H(s)$ presented in [Eq. @eq-PLL-closed-loop-transfer-function-initial] applies both to the phase of the feedback signal relative to the phase of the reference signal and to the frequency of the feedback signal relative to the frequency of the reference signal.

## Parameters and Components

The system parameters can be approximately calculated as follows:

$$
\omega_n = \sqrt{\frac{K}{C_1}}
$$ {#eq-omega_n-definition-no-C2}

and

$$
\zeta = \frac{R}{2}\sqrt{KC_1}
$$ {#eq-zeta-definition-no-C2}

Thus the components of the loop filter can be obtained as:

$$
C_1 = \frac{K}{\omega_n^2}
$$ {#eq-C1-loop-filter}

and:

$$
R_1 = \zeta\sqrt{\frac{2\omega_n}{K}}
$$ {#eq-R1-loop-filter}

For simplicity, the second capacitor can be obtained by $C_2=C_1/10$.

## Phase Error Transfer Function

The transfer function of the phase error relative to the reference phase can be obtained as:

$$
E(s)=\frac{\theta_e}{\theta_R}=H(s)\cdot\frac{s}{KF(s)}
$$ {#eq-tf-phase-error}

The example whose the resolution is being given here is studying the effect of an input frequency step. Since the input is described by phase, as it's possible to be seen in [Eq. @eq-phase-error-transfer-function-initial] ($\theta_R$), it's possible to take the phase equivalent of a frequency step, which is a phase ramp, since the phase is the integral of frequency, and the integral of a step is a ramp. Thus, the input can be described as:

$$
\theta_R=\frac{\Delta\omega}{s^2}
$$ {#eq-phase-of-frequency-step}

Getting the inverse Laplace transform of the respose of phase error transfer function for a step frequency, for $\zeta<1$, results in the following time-domain equation:

$$
\theta_e(t)=\frac{\Delta\omega}{\omega_n}\frac{\sin\left(\omega_n\sqrt{1-\zeta^2}t\right)}{\sqrt{1-\zeta^2}}e^{-\zeta\omega_nt}
$$ {#eq-inverse-laplace-phase-error}

### Maximum Step Frequency

At this point, the objective is to obtain an equation that provides the maximum frequency step the PLL can handle. To do so, the phase error equation is analyzed, and the peak value of the phase error is sought. Analyzing the maximum of [Eq. @eq-inverse-laplace-phase-error], the time instant at which the maximum phase error occurs, $t_m$, can be computed using:

$$
t_m=\frac{1}{\omega_n\sqrt{1-\zeta^2}}\tan^{-1}\left(\frac{\sqrt{1-\zeta^2}}{\zeta}\right)
$$ {#eq-instant-maximum-frequency-step}

[Eq. @eq-instant-maximum-frequency-step] represents the instant at which [Eq. @eq-inverse-laplace-phase-error] reaches its maximum value. Where the maximum normalized phase error is:

$$
\begin{split}
\theta_{e_\text{m,n}}&=\frac{\sin\left(\omega_n\sqrt{1-\zeta^2}t_m\right)}{\sqrt{1-\zeta^2}}e^{-\zeta\omega_nt_m}
\end{split}
$$

Thus:

$$
\theta_{e_m}=\frac{\Delta\omega_m}{\omega_n}\theta_{e_\text{m,n}}
$$

Leading to a maximum step frequency equal to:

$$
\Delta\omega_m=\frac{\theta_{e_m}\omega_n}{\theta_{e_\text{m,n}}}
$$ {#eq-maximum-step-frequency}

### Settling Time Estimation

The settling time, $t_s$, required for the system to reach a steady state after a frequency step can be approximated by:

$$
t_s\approx\frac{-\ln\left(0.01\sqrt{1-\zeta^2}\right)}{\zeta}\frac{1}{\omega_n}
$$

## Voltage Control Transfer Function

The transfer function of the control voltage relative to the reference phase can be obtained as:

$$
V_C(s)=\frac{V_C}{\theta_R}=H(s)\cdot\frac{1}{1/N}\cdot\frac{1}{K_\mathrm{VCO}}
$$ {#eq-tf-vctrl-vco}

## Output phase Transfer Function

The transfer function of the output phase relative to the reference phase can be obtained as:

$$
O(s)=\frac{V_C}{\theta_R}=H(s)\cdot\frac{1}{1/N}
$$ {#eq-tf-output-phase}

## Loop Natural Frequency Calculation

To compute the frequency bandwidth of the loop, the following equation is used:

$$
\omega_{3\text{dB}}=\omega_n\sqrt{1+2\zeta^2+\sqrt{4\zeta^4+4\zeta^2+2}}
$$ {#eq-frequency-bandwidth}

Rearanging the [Eq. @eq-frequency-bandwidth], it's possible to compute the natural frequency in terms of frequency bandwidth:

$$
\omega_n=\frac{\omega_{3\text{dB}}}{\sqrt{1+2\zeta^2+\sqrt{4\zeta^4+4\zeta^2+2}}}
$$ {#eq-natural-frequency-from-bandwidth}

## Cycle Slipping

Cycle slipping is an undesired phenomenon in Phase-Locked Loops (PLLs) that occurs when the phase error between the reference input and the feedback output exceeds the linear range of the system.

Cycle slipping occurs when the PLL is unable to maintain phase lock, and the phase error grows beyond $2\pi$, resulting in a temporary loss of synchronization. The loop then slips by one or more full cycles before attempting to reacquire lock.

## Validade da teoria

To ensure the theoretical models are valid, the minimum frequency of the reference signal must be respected. A commonly cited safe ratio is 10:1 for $\omega_r / \omega_n$. However, the minimum allowed frequency is given by:

$$
\frac{\omega_r}{\omega_n}\ge 2\pi\zeta
$$

If this rule is not followed, the linear models cannot be guaranteed, requiring that signal sampling is also taken into account and the system be treated as a discrete system, as it truly is.

# Design Analysis

```{python}
#| echo: false
import math

omega_n=omega_bw/math.sqrt(1+2*zeta**2+math.sqrt(4*zeta**4+4*zeta**2+2))
```

Based on the parameters already presented in the introduction, the natural frequency is $\omega_n = $ `{python} format_with_pi_latex(omega_n)`rad/s.

```{python}
#| echo: false
import math

t_m=1/(omega_n*math.sqrt(1-zeta**2))*math.atan(math.sqrt(1-zeta**2)/zeta)
```

With the parameters $\zeta$ and $\omega_n$ defined, it is possible to calculate the instant at which the maximum phase difference occurs. Thus, $t_m =$ `{python} format(t_m)`s.

```{python}
#| echo: false
import math

theta_e_m_n = math.sin(omega_n*math.sqrt(1-zeta**2)*t_m)/(math.sqrt(1-zeta**2))*math.e**(-zeta*omega_n*t_m)
theta_e_m = (1.6255) * math.pi
```

And applying $t_m$ to this equation, the maximum normalized phase error in this case is approximately `{python} round(theta_e_m_n, 2)`.

Normally, the maximum phase error supported by the PFD is expected to be $2\pi$. However, to match the values obtained with the simulator used, the maximum phase error is empirically determined as $\theta_{e_m} =$ `{python} format_with_pi_latex(theta_e_m)`.

```{python}
#| echo: false
import math

step_m_ang=theta_e_m*omega_n/theta_e_m_n
step_m_lin = step_m_ang/2/math.pi
```

Substituting the calculated values na [Eq. @eq-maximum-step-frequency] yields $\Delta\omega_m=$ `{python} format_with_pi_latex(step_m_ang)`rad/s or, alternativelly, $\Delta f_m=$ `{python} format(step_m_lin)`Hz. Hence, the maximum input frequency step that still allows theoretical predictability is `{python} format(step_m_lin)`Hz. If the input frequency step exceeds this limit, the PLL may lose lock and exhibit cycle slipping (more details are provided in the next section).

```{python}
#| echo: false
import math

t_s=math.log(0.01*math.sqrt(1-zeta**2), math.e)/(zeta*omega_n)*(-1)
```

With this frequency step and all parameters calculated so far, it is possible to compute the 1% settling time, resulting in $t_s =$ `{python} format(t_s)`s.

## Project

```{python}
#| echo: false

import numpy as np

# Parâmetros iniciais
f_r = 10e6 # Hz
omega_r = 2 * np.pi * f_r # rad/s
f_o = 50e6 # Hz
omega_o = 2 * np.pi * f_r # rad/s
N = 5
Kvco = KVCO = 10e6 # Hz/V
Kvco_ang = 2 * np.pi * Kvco # rad/s/V
I = 1e-3
K_theta = Kphase = I / (2 * np.pi) # A/rad
A0 = 6.258
K = A0 * Kphase * KVCO / N
```

To illustrate the results obtained so far, an integer-N Phase-Locked Loop (PLL) was developed with a division factor of $N =$ `{python} N`. The system was designed with a VCO gain of $K_\mathrm{VCO} =$ `{python} format(KVCO)`Hz/V, and a charge pump current of $I =$ `{python} format(I)`A. The phase detector gain is given by $K_\theta = I / 2\pi$. Furthermore, to align the calculated values with the responses obtained in simulation, an empirical value for the DC gain was determined as $A_0 =$ `{python} round(A0, 2)`.

```{python}
#| echo: false

from scipy import signal

omega_n = omega_bw / np.sqrt(1 + 2 * zeta**2 + np.sqrt(4 * zeta**4 + 4 * zeta**2 + 2))
f_n = omega_n/2/np.pi
```

The natural frequency of the loop, $\omega_n$, was computed from the loop bandwidth $\omega_\mathrm{bw}$ and damping factor $\zeta$, yielding a natural frequency $f_n=\omega_n/2\pi=$ `{python} format(f_n)`Hz.

```{python}
#| echo: false

# Parâmetros do PLL
C1 = K / omega_n**2
C2 = C1 / 10
R1 = 2 * zeta * omega_n / K
```

Based on the chosen parameters, the loop filter components were calculated as follows. The primary capacitor of the loop filter is given by the [Eq. @], that is, $C_1=$ `{python} format(C1)`F. The second capacitor is chosen to be one-tenth of $C_1$, i.e., $C_2=$ `{python} format(C2)`F. The resistor of the loop filter is given by $R_1=$ `{python} format(R1)`$\Omega$.

```{python}
#| echo: false

import math
import numpy as np

# Análise de resposta em frequência
omega_n_meas = np.sqrt(K / C1)
f_n_meas = omega_n_meas / (2 * np.pi)

zeta_meas = (R1 / 2) * np.sqrt(K * C1)

t_m_meas = 1 / (omega_n_meas * math.sqrt(1 - zeta_meas**2)) * math.atan(math.sqrt(1 - zeta_meas**2) / zeta_meas)

theta_e_m_n_meas = math.sin(omega_n_meas * math.sqrt(1 - zeta_meas**2) * t_m_meas) / \
                   (math.sqrt(1 - zeta_meas**2)) * math.e**(-zeta_meas * omega_n_meas * t_m_meas)

theta_e_m_meas = theta_e_m

omega_bw_meas = omega_n_meas * np.sqrt(1 + 2 * zeta_meas**2 + np.sqrt(4 * zeta_meas**4 + 4 * zeta_meas**2 + 2))
f_bw_meas = omega_bw_meas / (2 * np.pi)

step_m_ang_meas = theta_e_m_meas * omega_n_meas / theta_e_m_n_meas
step_m_lin_meas = step_m_ang_meas / (2 * math.pi)

t_s_meas = math.log(0.01 * math.sqrt(1 - zeta_meas**2)) / (zeta_meas * omega_n_meas) * (-1)

```

These component values define the loop filter required to achieve the desired dynamic behavior of the PLL. With these components, it's possible to extract another set of the system parameters, i.e., $f_n=\omega_n/2\pi$, $\zeta$ and $f_\mathrm{bw}=\omega_\mathrm{bw}/2\pi$. The values are $f_n=$ `{python} format(f_n_meas)`Hz, $\zeta=$ `{python} round(zeta_meas, 3)` and $f_\mathrm{bw}=$ `{python} format(f_bw_meas)`Hz.

```{python}
#| echo: false

from scipy import signal
from IPython.display import display, Markdown

def format_eng(x):
    if x == 0:
        return "0"
    exponent = int("{:.0e}".format(x).split('e')[1])
    base = x / (10**exponent)
    return f"{base:.2f}e{exponent}"

def poly_to_str(coeffs):
    terms = []
    degree = len(coeffs) - 1
    for i, coef in enumerate(coeffs):
        power = degree - i
        if abs(coef) < 1e-12:  # ignora coeficientes muito pequenos
            continue
        term = ""
        coef_str = format_eng(coef)
        if not (abs(coef - 1) < 1e-12 and power != 0):
            term += coef_str
        if power > 0:
            term += "s"
            if power > 1:
                term += f"^{power}"
        terms.append(term)
    return " + ".join(terms)
```

As shown in [Eq. @eq-tf-filter], the transfer function of the filter is given by:

```{python}
#| echo: false

# Numerador e denominador do compensador F(s)
# F(s) = (s*C1*R1 + 1) / (s^2*C1*C2*R1 + s*(C1 + C2))
F_num = [C1 * R1, 1]
F_den = [C1 * C2 * R1, C1 + C2, 0]

F_sys = signal.TransferFunction(F_num, F_den)
```

```{python}
#| echo: false
# Em scipy.signal, num e den são arrays dentro de listas (mesmo para SISO)
num = F_sys.num[0][0] if isinstance(F_sys.num[0], (list, tuple)) else F_sys.num
den = F_sys.den[0][0] if isinstance(F_sys.den[0], (list, tuple)) else F_sys.den

num_str = poly_to_str(num)
den_str = poly_to_str(den)

# Exibe como equação no Jupyter
display(Markdown(f"$$F(s) = \\frac{{{num_str}}}{{{den_str}}}$$ {{#eq-tf-filter-numeric}}"))

```

The open-loop transfer function of the system is defined in [Eq. @eq-tf-open-loop], as follows:

```{python}
#| echo: false
# F(s) * K / s = G(s)
G_num = np.polymul(F_num, [K])
G_den = np.polymul(F_den, [1, 0])  # Multiplica por s

G_sys = signal.TransferFunction(G_num, G_den)
```

```{python}
#| echo: false
# Em scipy.signal, num e den são arrays dentro de listas (mesmo para SISO)
num = G_sys.num[0][0] if isinstance(G_sys.num[0], (list, tuple)) else G_sys.num
den = G_sys.den[0][0] if isinstance(G_sys.den[0], (list, tuple)) else G_sys.den

num_str = poly_to_str(num)
den_str = poly_to_str(den)

# Exibe como equação no Jupyter
display(Markdown(f"$$G(s) = \\frac{{{num_str}}}{{{den_str}}}$$ {{#eq-tf-open-loop-numeric}}"))
```

As illustrated in [Eq. @eq-tf-closed-loop], the closed-loop transfer function of the system can be expressed as:

```{python}
#| echo: false
# H(s) = G(s) / (1 + G(s))
H_num = G_num
H_den = np.polyadd(G_den, G_num)

# Criando o sistema de tempo contínuo
H_sys = signal.TransferFunction(H_num, H_den)
```

```{python}
#| echo: false
# Em scipy.signal, num e den são arrays dentro de listas (mesmo para SISO)
num = H_sys.num[0][0] if isinstance(H_sys.num[0], (list, tuple)) else H_sys.num
den = H_sys.den[0][0] if isinstance(H_sys.den[0], (list, tuple)) else H_sys.den

num_str = poly_to_str(num)
den_str = poly_to_str(den)

# Exibe como equação no Jupyter
display(Markdown(f"$$H(s) = \\frac{{{num_str}}}{{{den_str}}}$$ {{#eq-tf-closed-loop-numeric}}"))
```

The transfer function of the output phase is provided in [Eq. @eq-tf-output-phase], as follows:

```{python}
#| echo: false

# Agora temos o denominador completo
# O_sys = H_sys * N
# Para isso, O_sys = (H_num * N) / (H_den)
num_O = np.convolve(H_sys.num, [N])
den_O = np.convolve(H_sys.den, [1])

# Criar a função de transferência final
O_sys = signal.TransferFunction(num_O, den_O)
```

```{python}
#| echo: false
# Em scipy.signal, num e den são arrays dentro de listas (mesmo para SISO)
num = O_sys.num[0][0] if isinstance(O_sys.num[0], (list, tuple)) else O_sys.num
den = O_sys.den[0][0] if isinstance(O_sys.den[0], (list, tuple)) else O_sys.den

num_str = poly_to_str(num)
den_str = poly_to_str(den)

# Exibe como equação no Jupyter
display(Markdown(f"$$O(s) = \\frac{{{num_str}}}{{{den_str}}}$$ {{#eq-tf-output-phase-numeric}}"))
```

The transfer function of the phase error is presented in [Eq. @eq-tf-phase-error], described by:

```{python}
#| echo: false

# Kvco/s => [Kvco] / [1, 0]
Kvco_over_s = ([Kvco], [1, 0])

# F_sys => ([1], [1, 1]) já está
# Multiplicando numeradores e denominadores via convolução
num_den = np.convolve([1/N], Kvco_over_s[0])
num_den = np.convolve(num_den, F_sys.num)
num_den = np.convolve(num_den, [K_theta])
num_den = np.convolve(num_den, [A0])

den_den = np.convolve(Kvco_over_s[1], F_sys.den)

# Agora temos o denominador completo
# E_sys = H_sys / (den)
# Para isso, E_sys = (H_num * den_den) / (H_den * den_num)
num_E = np.convolve(H_sys.num, den_den)
den_E = np.convolve(H_sys.den, num_den)

# Criar a função de transferência final
E_sys = signal.TransferFunction(num_E, den_E)
```

```{python}
#| echo: false
# Em scipy.signal, num e den são arrays dentro de listas (mesmo para SISO)
num = E_sys.num[0][0] if isinstance(E_sys.num[0], (list, tuple)) else E_sys.num
den = E_sys.den[0][0] if isinstance(E_sys.den[0], (list, tuple)) else E_sys.den

num_str = poly_to_str(num)
den_str = poly_to_str(den)

# Exibe como equação no Jupyter
display(Markdown(f"$$E(s) = \\frac{{{num_str}}}{{{den_str}}}$$ {{#eq-tf-phase-error-numeric}}"))
```

As detailed in [Eq. @eq-tf-vctrl-vco], the transfer function of the control voltage of the VCO is given by:

```{python}
#| echo: false

# Vc_sys = H_sys / ((1/N)*(Kvco/s))
#Kvco_over_s = ([Kvco], [1, 0])
Kvco_over_s = ([Kvco], [1])
den_num = np.convolve([1/N], Kvco_over_s[0])
den_den = Kvco_over_s[1]

num_Vc = np.convolve(H_sys.num, den_den)
den_Vc = np.convolve(H_sys.den, den_num)

Vc_sys = signal.TransferFunction(num_Vc, den_Vc)
```

```{python}
#| echo: false
# Em scipy.signal, num e den são arrays dentro de listas (mesmo para SISO)
num = Vc_sys.num[0][0] if isinstance(Vc_sys.num[0], (list, tuple)) else Vc_sys.num
den = Vc_sys.den[0][0] if isinstance(Vc_sys.den[0], (list, tuple)) else Vc_sys.den

num_str = poly_to_str(num)
den_str = poly_to_str(den)

# Exibe como equação no Jupyter
display(Markdown(f"$$V_C(s) = \\frac{{{num_str}}}{{{den_str}}}$$ {{#eq-tf-vctrl-vco-numeric}}"))
```

## Validade do modelo

```{python}
#| echo: false

import numpy as np

omega_r_safe = omega_n*10
f_r_safe = omega_r_safe/2/np.pi

omega_r_max = omega_n * 2 * np.pi * zeta
f_r_max = omega_r_max/2/np.pi
```

The minimum safe frequency for the reference signal is given by $f_r=$ `{python} format(f_r_safe)`Hz. However, the minimum reference frequency is $f_r\ge$ `{python} format(f_r_max)`Hz. Therefore, since a PLL operating with a reference frequency of $f_r=$ `{python} format(f_r)`Hz and an output frequency of `{python} format(f_o)`Hz is desired, the designed system must operate correctly.

# System Responses

```{python}
#| echo: false

sim_stop_time = 50e-6
sim_points = 10000
step_amplitude = f_r
```

As shown in [Fig. @fig-step-freq-h], a frequency step input of `{python} format(step_amplitude)`Hz is applied to evaluate the dynamic behavior of the closed-loop system described by the transfer function in [Eq. @eq-tf-closed-loop-numeric]. This analysis reveals the system’s capacity to reach a new steady-state condition in response to abrupt input changes and the settling time.

```{python}
#| echo: false
#| label: fig-step-freq-h
#| fig-cap: "Frequency step response of the closed-loop transfer function."
#| fig-align: center

import numpy as np
import matplotlib.pyplot as plt
from scipy import signal

# Tempo de simulação
t = np.linspace(0, sim_stop_time, sim_points)

# Criar vetor da entrada (degrau de amplitude 5)
u = np.ones_like(t) * step_amplitude

# Simulação da resposta ao degrau via lsim
t_out, y_out, _ = signal.lsim(H_sys, U=u, T=t)

# Valor final da resposta
final_value = y_out[-1]
tol = 0.01 * final_value  # 1% da resposta final

# Encontrar o tempo em que a resposta entra e permanece na faixa de ±1%
ts_graphical = 0
within_bounds = np.abs(y_out - u) <= tol
for i in range(len(within_bounds)):
    if all(within_bounds[i:]):
        ts_graphical = t_out[i]
        break

# Gerar gráfico
plt.figure(figsize=(8, 6))
plt.plot(t_out * 1e6, y_out, label='Feedback frequency')
plt.plot(t_out * 1e6, u, linestyle=':', color='black', label='Reference frequency')
plt.axhline(final_value + tol, color='red', linestyle='--', linewidth=0.8, label='±1% Tolerance')
plt.axhline(final_value - tol, color='red', linestyle='--', linewidth=0.8)
plt.axvline(ts_graphical * 1e6, color='green', linestyle='--', linewidth=1.2, label=f'Settling time: {ts_graphical*1e6:.5f} µs')
plt.xlabel('Time (µs)')
plt.ylabel('Frequency (Hz)')
plt.grid(True)
plt.legend()
plt.tight_layout()
plt.show()
```

As illustrated in [Fig. @fig-ramp-phase-h], a phase ramp input is employed to assess the tracking capabilities of the closed-loop system represented by [Eq. @eq-tf-closed-loop]. This test simulates a constant frequency offset and allows for the evaluation of the system’s ability to maintain phase lock.

```{python}
#| echo: false
#| label: fig-ramp-phase-h
#| fig-cap: "Phase response to a ramp input in the closed-loop transfer function."
#| fig-align: center

import numpy as np
import matplotlib.pyplot as plt
from scipy import signal

# Tempo de simulação
t = np.linspace(0, sim_stop_time, sim_points)

# Criar vetor da entrada (degrau de amplitude 5)
step_amplitude = omega_r * t
u = np.ones_like(t) * step_amplitude

# Simulação da resposta ao degrau via lsim
t_out, y_out, _ = signal.lsim(H_sys, U=u, T=t)

# Gerar gráfico
plt.figure(figsize=(8, 6))
plt.plot(t_out * 1e6, y_out, label='Feedback Phase')
plt.plot(t_out * 1e6, u, linestyle=':', color='black', label='Reference Phase')
plt.xlabel('Time (µs)')
plt.ylabel('Phase (rad)')
plt.grid(True)
plt.legend()
plt.tight_layout()
plt.show()
```

[Fig. @fig-ramp-phase-h-o] presents a comparative analysis of the phase responses of the closed-loop and output-phase transfer functions, as defined in [Eq. @eq-tf-closed-loop] and [Eq. @eq-tf-output-phase], respectively. The figure highlights their individual contributions to the system’s overall phase behavior under ramp input conditions.

```{python}
#| echo: false
#| label: fig-ramp-phase-h-o
#| fig-cap: "Comparison of phase responses to a ramp input in two different functions."
#| fig-align: center

import numpy as np
import matplotlib.pyplot as plt
from scipy import signal

# Tempo de simulação
t = np.linspace(0, sim_stop_time, sim_points)

# Criar vetor da entrada (degrau de amplitude 5)
step_amplitude = omega_r * t
u = np.ones_like(t) * step_amplitude

# Simulação da resposta ao degrau via lsim
t_out, y_out, _ = signal.lsim(H_sys, U=u, T=t)
t_out, y_out2, _ = signal.lsim(O_sys, U=u, T=t)

# Gerar gráfico
plt.figure(figsize=(8, 6))
plt.plot(t_out * 1e6, y_out, label='Feedback Phase')
plt.plot(t_out * 1e6, y_out2, label='Output Phase')
plt.plot(t_out * 1e6, u, linestyle=':', color='black', label='Reference Phase')
plt.xlabel('Time (µs)')
plt.ylabel('Phase (rad)')
plt.grid(True)
plt.legend()
plt.tight_layout()
plt.show()
```

In [Fig. @fig-ramp-phase-step-freq-h], a ramp input with a superimposed frequency step is used to analyze the transient behavior of the closed-loop system, whose response is characterized by [Eq. @eq-tf-closed-loop]. This approach reveals how the system adapts to dynamic input variations.

```{python}
#| echo: false
#| label: fig-ramp-phase-step-freq-h
#| fig-cap: "Phase response to a ramp with an abrupt frequency change (frequency step)."
#| fig-align: center

import numpy as np
import matplotlib.pyplot as plt
from scipy import signal

# Tempo de simulação
t = np.linspace(0, sim_stop_time, sim_points)

# Frequências
f_base = omega_r  # 10 MHz
f_step = step_m_ang_meas  # step de 10 MHz
t_step = 20e-6  # instante do step

# Entrada em fase: integral da frequência
phi_in = np.where(
    t < t_step,
    f_base * t,
    f_base * t + f_step * (t - t_step)
)

# Simulação da resposta
t_out, y_out, _ = signal.lsim(H_sys, U=phi_in, T=t)

### --- Gráfico ---
plt.figure(figsize=(8, 6))
plt.plot(t_out * 1e6, y_out, label='Feedback Phase')
plt.plot(t_out * 1e6, phi_in, linestyle=':', color='black', label='Reference Phase')

plt.axvline(t_step * 1e6, color='purple', linestyle='--', linewidth=1.2, label='Step Frequency Instant')

plt.xlabel('Time (µs)')
plt.ylabel('Phase (rad)')
plt.grid(True)
plt.legend()
plt.tight_layout()
plt.show()

```

As depicted in [Fig. @fig-ramp-phase-h-phase-error], the response of the closed-loop system under a ramp input is evaluated with particular attention to the resulting phase error. The system behavior follows the model described by [Eq. @eq-tf-closed-loop], and this evaluation helps quantify the residual phase deviation.

```{python}
#| echo: false
#| label: fig-ramp-phase-h-phase-error
#| fig-cap: "Response to a ramp input with phase error."
#| fig-align: center

import numpy as np
import matplotlib.pyplot as plt
from scipy import signal

# Tempo de simulação
t = np.linspace(0, sim_stop_time, sim_points)

# Entrada: rampa em fase (step de 10 MHz em frequência)
# Fase = ∫ frequência ⇒ rampa
freq_step = omega_r  # Step de 10 MHz em frequência
phi_in = freq_step * t  # Rampa de fase

# Simulação da resposta do sistema à rampa
t_out, phi_out, _ = signal.lsim(H_sys, U=phi_in, T=t)

# Erro de seguimento
tracking_error = phi_in - phi_out

# Plotagem
plt.figure(figsize=(8, 6))
plt.plot(t_out * 1e6, phi_in, linestyle=':', color='black', label='Reference Phase')
plt.plot(t_out * 1e6, phi_out, label='Feedback Phase')
plt.plot(t_out * 1e6, tracking_error, label='Phase Error')
plt.xlabel('Time (µs)')
plt.ylabel('Phase (rad)')
plt.grid(True)
plt.legend()
plt.tight_layout()
plt.show()

```

The evolution of the phase error in response to a frequency step input is shown in [Fig. @fig-step-freq-e], and it corresponds to the transfer function presented in [Eq. @eq-tf-phase-error]. This analysis highlights the loop's capacity to promptly correct for frequency disturbances.

```{python}
#| echo: false
#| label: fig-step-freq-e
#| fig-cap: "Phase error response to a frequency step input."
#| fig-align: center

import numpy as np
import matplotlib.pyplot as plt
from scipy import signal

# Tempo de simulação
t = np.linspace(0, sim_stop_time, sim_points)

# Criar vetor da entrada (degrau de amplitude 5)
step_amplitude = step_m_ang_meas * t
u = np.ones_like(t) * step_amplitude

# Simulação da resposta ao degrau via lsim
t_out, y_out, _ = signal.lsim(E_sys, U=u, T=t)

# Cálculo do valor máximo (pico)
max_value = np.max(np.abs(y_out))
tol = 0.01 * max_value  # 1% do valor de pico (centrado em zero)

# Novo cálculo do tempo de acomodação (settling time)
within_bounds = np.abs(y_out) <= tol
ts_graphical_peak = 0
for i in range(len(within_bounds)):
    if all(within_bounds[i:]):
        ts_graphical_peak = t_out[i]
        break

# Gerar gráfico
plt.figure(figsize=(8, 6))
plt.plot(t_out * 1e6, y_out, label='Phase Error')

# Linha no valor máximo (positivo)
plt.axhline(max_value, color='blue', linestyle=':', linewidth=1.2, label=f'Max Value')

# Linhas horizontais em ±1% do valor de pico (em torno de zero)
plt.axhline(tol, color='red', linestyle='--', linewidth=0.8, label=f'±1% Tolerance')
plt.axhline(-tol, color='red', linestyle='--', linewidth=0.8)

# Marcação do novo tempo de acomodação
plt.axvline(ts_graphical_peak * 1e6, color='green', linestyle='--', linewidth=1.2, label=f'Settling Time: {ts_graphical_peak*1e6:.3f} μs')

plt.xlabel('Time (µs)')
plt.ylabel('Phase Error (rad)')
plt.grid(True)
plt.legend()
plt.tight_layout()
plt.show()

```

[Fig. @fig-step-freq-e-norm] illustrates the normalized version of the phase error response, which is governed by the transfer function in [Eq. @eq-tf-phase-error]. Normalization enables meaningful comparisons across different system designs.

```{python}
#| echo: false
#| label: fig-step-freq-e-norm
#| fig-cap: "Normalized phase error response to a frequency step input."
#| fig-align: center

import numpy as np
import matplotlib.pyplot as plt
from scipy import signal

# Tempo de simulação
t = np.linspace(0, 8/omega_n_meas, sim_points)

# Criar vetor da entrada (degrau de amplitude 5)
step_amplitude = t
u = np.ones_like(t) * step_amplitude

# Simulação da resposta ao degrau via lsim
t_out, y_out, _ = signal.lsim(E_sys, U=u, T=t)

y_out *= omega_n_meas

# Cálculo do valor máximo (pico)
max_value = np.max(np.abs(y_out))
tol = 0

# Gerar gráfico
plt.figure(figsize=(8, 6))
plt.plot(t_out * omega_n_meas, y_out, label='Step Response')

# Linha no valor máximo (positivo)
plt.axhline(max_value, color='blue', linestyle=':', linewidth=1.2, label=f'Max Value: {max_value:.3f}')

# Linhas horizontais em ±1% do valor de pico (em torno de zero)
plt.axhline(tol, color='black', linestyle=':')

plt.xlabel('Time (µs)')
plt.ylabel('Phase Error norm. (rad)')
plt.grid(True)
plt.legend()
plt.tight_layout()
plt.show()

```

As can be observed in [Fig. @fig-step-freq-vc], the control voltage applied to the VCO is analyzed following a step in input frequency. The dynamic behavior is described by the transfer function in [Eq. @eq-tf-vctrl-vco], which reflects the loop's ability to adjust its frequency through the VCO.

```{python}
#| echo: false
#| label: fig-step-freq-vc
#| fig-cap: "Control voltage response to a frequency step input."
#| fig-align: center

import numpy as np
import matplotlib.pyplot as plt
from scipy.signal import TransferFunction, step

# Tempo de simulação
t = np.linspace(0, sim_stop_time, sim_points)

# Resposta ao degrau
t, vc_response = step(Vc_sys, T=t)
vc_response *= step_m_lin_meas

# Valor final (assumido ou estimado)
final_value = ((f_r+step_m_lin) * N - f_o) / KVCO
#final_value = 0.5
tol = 0.01 * final_value  # 1% de tolerância
upper_bound = final_value + tol
lower_bound = final_value - tol

# Encontrar settling time: busca o primeiro instante em que o sinal entra e permanece dentro da faixa de tolerância
within_bounds = np.logical_and(vc_response >= lower_bound, vc_response <= upper_bound)

# Janela de amostras para verificar se permanece dentro da tolerância
window_size = 3*150  # ~5% da simulação

ts_graphical = 0
for i in range(len(within_bounds) - window_size):
    if np.all(within_bounds[i:i + window_size]):
        ts_graphical = t[i]
        break

# Plotagem
plt.figure(figsize=(8, 6))
plt.plot(t * 1e6, vc_response, label='Voltage Control')
plt.axhline(upper_bound, color='red', linestyle='--', linewidth=0.8, label='±1% Tolerance')
plt.axhline(lower_bound, color='red', linestyle='--', linewidth=0.8)

if ts_graphical is not None:
    plt.axvline(ts_graphical * 1e6, color='green', linestyle='--', linewidth=1.2,
                label=f'Settling time: {ts_graphical*1e6:.2f} µs')

plt.xlabel('Time (µs)')
plt.ylabel('Voltage Control (V)')
plt.grid(True)
plt.legend()
plt.tight_layout()
plt.show()

```

[Fig. @fig-step-freq-vc-norm] presents the normalized control voltage response governed by [Eq. @eq-tf-vctrl-vco]. The normalized plot facilitates the comparison of different control loop designs by removing absolute voltage scale dependencies.

```{python}
#| echo: false
#| label: fig-step-freq-vc-norm
#| fig-cap: "Normalized control voltage response to a frequency step input."
#| fig-align: center

import numpy as np
import matplotlib.pyplot as plt
from scipy.signal import TransferFunction, step

# Tempo de simulação
t = np.linspace(0, 8/omega_n_meas, sim_points)

# Resposta ao degrau
t, vc_response = step(Vc_sys, T=t)
vc_response *= Kvco / N

# Valor final (assumido ou estimado)
final_value = ((f_r+step_m_lin) * N - f_o) / KVCO
final_value *= KVCO / (step_m_lin_meas * N)
#final_value = 0.5

# Janela de amostras para verificar se permanece dentro da tolerância
window_size = 50  # ~5% da simulação

# Plotagem
plt.figure(figsize=(8, 6))
plt.plot(t * omega_n_meas, vc_response, label='Voltage Control')
plt.axhline(final_value, color='black', linestyle=':', label='Reference Voltage')

plt.xlabel('Tempo (µs)')
plt.ylabel('Voltage Control norm. (V)')
plt.grid(True)
plt.legend()
plt.tight_layout()
plt.show()

```

As illustrated in the Bode diagram shown in [Fig. @fig-bode-g], the open-loop transfer function of the system, defined in [Eq. @eq-tf-open-loop], characterizes the frequency response in terms of gain and phase. This representation enables the analysis of system stability and the determination of key dynamic properties such as gain margin and phase margin.

```{python}
#| echo: false
#| label: fig-bode-g
#| fig-cap: "Bode plot of the open-loop transfer function."
#| fig-align: center

import numpy as np
import matplotlib.pyplot as plt
from scipy import signal

# Gerar a resposta em frequência
w, mag, phase = signal.bode(G_sys)

# Plotando o diagrama de Bode
plt.figure(figsize=(8, 6))

# Magnitude
plt.subplot(2, 1, 1)
plt.semilogx(w, mag, label='Magnitude (dB)')
plt.ylabel('Magnitude (dB)')
plt.grid(True, which="both", ls="--")
plt.legend()

# Fase
plt.subplot(2, 1, 2)
plt.semilogx(w, phase)
plt.ylabel('Phase (degrees)')
plt.xlabel('Frequency (rad/s)')
plt.grid(True, which="both", ls="--")

plt.tight_layout()
plt.show()
```

The frequency-domain characteristics of the closed-loop system, modeled by [Eq. @eq-tf-closed-loop], are depicted in [Fig. @fig-bode-h]. This Bode plot provides insight into stability margins, bandwidth, and overall system dynamics.

```{python}
#| echo: false
#| label: fig-bode-h
#| fig-cap: "Bode plot of the closed-loop transfer function."
#| fig-align: center

import numpy as np
import matplotlib.pyplot as plt
from scipy import signal

# Gerar a resposta em frequência
w, mag, phase = signal.bode(H_sys)

# Encontrar o ponto onde a magnitude cruza -3 dB
cutoff_index = np.argmax(mag <= -3)
cutoff_freq_ang = w[cutoff_index]
cutoff_freq_lin = cutoff_freq_ang/2/np.pi
cutoff_mag = mag[cutoff_index]

# Plotando o diagrama de Bode
plt.figure(figsize=(8, 6))

# Magnitude
plt.subplot(2, 1, 1)
plt.semilogx(w, mag, label='Magnitude (dB)')
plt.axvline(x=cutoff_freq_ang, color='r', linestyle='--', label='Cutoff Frequency (-3 dB)')
plt.axhline(y=-3, color='gray', linestyle=':', label='-3 dB')
plt.scatter(cutoff_freq_ang, cutoff_mag, color='red')  # Marcar o ponto

# Adiciona o valor da frequência no gráfico
plt.text(cutoff_freq_ang, cutoff_mag - 5, f'{format_graph(cutoff_freq_lin)}Hz', color='red',
         ha='center', va='top', fontsize=9, bbox=dict(facecolor='white', edgecolor='red', boxstyle='round'))

plt.ylabel('Magnitude (dB)')
plt.grid(True, which="both", ls="--")
plt.legend()

# Fase
plt.subplot(2, 1, 2)
plt.semilogx(w, phase)
plt.axvline(x=cutoff_freq_ang, color='r', linestyle='--')
plt.ylabel('Phase (degrees)')
plt.xlabel('Frequency (rad/s)')
plt.grid(True, which="both", ls="--")

plt.tight_layout()
plt.show()
```

In [Fig. @fig-bode-f], the gain and phase of the loop filter, as defined in [Eq. @eq-tf-filter], are plotted. This representation is key for understanding how the filter shapes the loop response and contributes to noise suppression and stability.

```{python}
#| echo: false
#| label: fig-bode-f
#| fig-cap: "Bode plot of the filter transfer function."
#| fig-align: center

import numpy as np
import matplotlib.pyplot as plt
from scipy import signal

# Gerar a resposta em frequência
w, mag, phase = signal.bode(F_sys)

# Encontrar o ponto onde a magnitude cruza -3 dB
cutoff_index = np.argmax(mag <= -3)
cutoff_freq_ang = w[cutoff_index]
cutoff_freq_lin = cutoff_freq_ang/2/np.pi
cutoff_mag = mag[cutoff_index]

# Plotando o diagrama de Bode
plt.figure(figsize=(8, 6))

# Magnitude
plt.subplot(2, 1, 1)
plt.semilogx(w, mag, label='Magnitude (dB)')

plt.ylabel('Magnitude (dB)')
plt.grid(True, which="both", ls="--")
plt.legend()

# Fase
plt.subplot(2, 1, 2)
plt.semilogx(w, phase)
plt.ylabel('Phase (degrees)')
plt.xlabel('Frequency (rad/s)')
plt.grid(True, which="both", ls="--")

plt.tight_layout()
plt.show()
```

# Conclusion

This report provides a theoretical analysis of a second-order PFD/CP-based Integer-N PLL. The maximum tolerable frequency step for which the linear model remains valid was derived, and the corresponding settling time was estimated. These results serve as a foundation for further simulation and validation studies.

# References {-}